---
title: "Colorado Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

party_shell_returned <- data.frame(Party=character(),
                                  Count=integer(),
                                  Frequency=double(),
                                  Count2=integer(),
                                  Rate=integer(),
                                  stringsAsFactors=FALSE)
party_shell_returned[1,1] <- "Democrats"
party_shell_returned[2,1] <- "Republicans"
party_shell_returned[3,1] <- "Minor"
party_shell_returned[4,1] <- "No Party Affiliation"
party_shell_returned[5,1] <- "TOTAL"



# Colorado

CO_req_send_party <- party_shell

CO_req_send_party[1,2] <- state_stats[6,10] 
CO_req_send_party[2,2] <- state_stats[6,11] 
CO_req_send_party[3,2] <- state_stats[6,12] 
CO_req_send_party[4,2] <- state_stats[6,13] 
CO_req_send_party[5,2] <- state_stats[6,5] 

CO_req_send_party$Percent <- 100*CO_req_send_party$Count/CO_req_send_party[5,2]

CO_accept_party <- party_shell_returned

CO_accept_party[1,2] <- state_stats[6,29] 
CO_accept_party[2,2] <- state_stats[6,30] 
CO_accept_party[3,2] <- state_stats[6,31] 
CO_accept_party[4,2] <- state_stats[6,32] 
CO_accept_party[5,2] <- state_stats[6,6] 

CO_accept_party[1,4] <- state_stats[6,10] 
CO_accept_party[2,4] <- state_stats[6,11] 
CO_accept_party[3,4] <- state_stats[6,12] 
CO_accept_party[4,4] <- state_stats[6,13] 
CO_accept_party[5,4] <- state_stats[6,5] 

CO_accept_party$Frequency <- 100*CO_accept_party$Count/CO_accept_party[5,2]
CO_accept_party$Rate <- 100*CO_accept_party$Count/CO_accept_party$Count2

colnames(CO_accept_party) <- c("Party", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")



CO_req_send_party <- party_shell

CO_req_send_party[1,2] <- state_stats[6,10] 
CO_req_send_party[2,2] <- state_stats[6,11] 
CO_req_send_party[3,2] <- state_stats[6,12] 
CO_req_send_party[4,2] <- state_stats[6,13] 
CO_req_send_party[5,2] <- state_stats[6,5] 

CO_req_send_party$Percent <- 100*CO_req_send_party$Count/CO_req_send_party[5,2]

CO_req_send_gender <- gender_shell

CO_req_send_gender[1,2] <- sum(state_stats[6,21])
CO_req_send_gender[2,2] <- sum(state_stats[6,20]) 
CO_req_send_gender[3,2] <- sum(state_stats[6,22]) 
CO_req_send_gender[4,2] <- sum(state_stats[6,5]) 

CO_req_send_gender$Percent <- 100 * CO_req_send_gender$Count/CO_req_send_gender[4,2]

CO_req_send_age <- age_shell

CO_req_send_age[1,2] <- state_stats[6,23] 
CO_req_send_age[2,2] <- state_stats[6,24] 
CO_req_send_age[3,2] <- state_stats[6,25] 
CO_req_send_age[4,2] <- state_stats[6,26] 
CO_req_send_age[5,2] <- state_stats[6,27] 
CO_req_send_age[6,2] <- state_stats[6,28] 
CO_req_send_age[7,2] <- state_stats[6,5] 

CO_req_send_age$Percent <- 100*CO_req_send_age$Count/CO_req_send_age[7,2]
```
## {.tabset}
Last Report: `r state_stats[6,9]`  
Source:  `r state_stats[6,2]`

Total Early Vote: **`r format(as.numeric(state_stats[6,8]), big.mark =",")`**  
Total Early Vote as Precentage of 2016 Total Turnout: **`r paste(round(100*as.integer(state_stats[6,8])/as.integer(state_stats[6,3]), digits = 1),"%", sep = "")`**

### Mail Ballots Returned

Ballots Returned: **`r format(as.numeric(state_stats[6,6]), big.mark =",")`**

#### **Mail Ballots Returned and Accepted by Party Registration**

``` {r echo = FALSE}
kable(CO_accept_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

### Requested Mail Ballots

Ballots Requested: **`r format(as.numeric(state_stats[6,5]), big.mark =",")`**

Colorado is an all-mail ballot state. Ballot requests reported here are counts of the state's registered voters.

#### Mail Ballot Requests by Party Registration

``` {r echo = FALSE}
kable(CO_req_send_party,format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

Colorado registered Democrats have a **`r format(as.numeric(CO_req_send_party[1,2]-CO_req_send_party[2,2]), big.mark =",")`** ballot request lead over registered Republicans.

#### Mail Ballot Requests by Age

``` {r echo = FALSE}
kable(CO_req_send_age, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

