---
title: "2020 General Election Early Vote Statistics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

# National

US_req_send_party <- party_shell

US_req_send_party[1,2] <- sum(state_stats$mail_sent_req_2020_dem) 
US_req_send_party[2,2] <- sum(state_stats$mail_sent_req_2020_rep) 
US_req_send_party[3,2] <- sum(state_stats$mail_sent_req_2020_minor)
US_req_send_party[4,2] <- sum(state_stats$mail_sent_req_2020_npa) 
US_req_send_party[5,2] <- sum(state_stats$mail_sent_req_2020_dem)+sum(state_stats$mail_sent_req_2020_rep)+sum(state_stats$mail_sent_req_2020_minor)+sum(state_stats$mail_sent_req_2020_npa)

US_req_send_party$Percent <- 100*US_req_send_party$Count/US_req_send_party[5,2]

US_req_send_race <- race_shell

US_req_send_race[1,2] <- sum(state_stats$mail_sent_req_2020_nh_white) 
US_req_send_race[2,2] <- sum(state_stats$mail_sent_req_2020_nh_black) 
US_req_send_race[3,2] <- sum(state_stats$mail_sent_req_2020_hispanic)
US_req_send_race[4,2] <- sum(state_stats$mail_sent_req_2020_nh_asian) 
US_req_send_race[5,2] <- sum(state_stats$mail_sent_req_2020_nh_native_american)
US_req_send_race[6,2] <- sum(state_stats$mail_sent_req_2020_other) 
US_req_send_race[7,2] <- sum(state_stats$mail_sent_req_2020_nh_white) +
  sum(state_stats$mail_sent_req_2020_nh_black) +
  sum(state_stats$mail_sent_req_2020_hispanic) +
  sum(state_stats$mail_sent_req_2020_nh_asian) +
  sum(state_stats$mail_sent_req_2020_nh_native_american) +
  sum(state_stats$mail_sent_req_2020_other)

US_req_send_race$Percent <- 100*US_req_send_race$Count/US_req_send_race[7,2]

US_req_send_age <- age_shell

US_req_send_age[1,2] <- sum(state_stats$mail_sent_req_2020_age1824) 
US_req_send_age[2,2] <- sum(state_stats$mail_sent_req_2020_age2534) 
US_req_send_age[3,2] <- sum(state_stats$mail_sent_req_2020_age3544)
US_req_send_age[4,2] <- sum(state_stats$mail_sent_req_2020_age4554)
US_req_send_age[5,2] <- sum(state_stats$mail_sent_req_2020_age5564)
US_req_send_age[6,2] <- sum(state_stats$mail_sent_req_2020_age65up)
US_req_send_age[7,2] <- sum(state_stats$mail_sent_req_2020_age1824) +
  sum(state_stats$mail_sent_req_2020_age2534) +
  sum(state_stats$mail_sent_req_2020_age3544) +
  sum(state_stats$mail_sent_req_2020_age4554) +
  sum(state_stats$mail_sent_req_2020_age5564) +
  sum(state_stats$mail_sent_req_2020_age65up)

US_req_send_age$Percent <- 100*US_req_send_age$Count/US_req_send_age[7,2]



US_req_send_gender <- gender_shell

US_req_send_gender[1,2] <- sum(state_stats$mail_sent_req_2020_female) 
US_req_send_gender[2,2] <- sum(state_stats$mail_sent_req_2020_male) 
US_req_send_gender[3,2] <- sum(state_stats$mail_sent_req_2020_unk)
US_req_send_gender[4,2] <- sum(state_stats$mail_sent_req_2020_female) +
  sum(state_stats$mail_sent_req_2020_male) +
  sum(state_stats$mail_sent_req_2020_unk)

US_req_send_gender$Percent <- 100*US_req_send_gender$Count/US_req_send_gender[4,2]

# Colorado

CO_req_send_party <- party_shell

CO_req_send_party[1,2] <- state_stats[6,10] 
CO_req_send_party[2,2] <- state_stats[6,11] 
CO_req_send_party[3,2] <- state_stats[6,12] 
CO_req_send_party[4,2] <- state_stats[6,13] 
CO_req_send_party[5,2] <- state_stats[6,5] 

CO_req_send_party$Percent <- 100*CO_req_send_party$Count/CO_req_send_party[5,2]

CO_req_send_gender <- gender_shell

CO_req_send_gender[1,2] <- sum(state_stats[6,21])
CO_req_send_gender[2,2] <- sum(state_stats[6,20]) 
CO_req_send_gender[3,2] <- sum(state_stats[6,22]) 
CO_req_send_gender[4,2] <- sum(state_stats[6,5]) 

CO_req_send_gender$Percent <- 100 * CO_req_send_gender$Count/CO_req_send_gender[4,2]

CO_req_send_age <- age_shell

CO_req_send_age[1,2] <- state_stats[6,23] 
CO_req_send_age[2,2] <- state_stats[6,24] 
CO_req_send_age[3,2] <- state_stats[6,25] 
CO_req_send_age[4,2] <- state_stats[6,26] 
CO_req_send_age[5,2] <- state_stats[6,27] 
CO_req_send_age[6,2] <- state_stats[6,28] 
CO_req_send_age[7,2] <- state_stats[6,5] 

CO_req_send_age$Percent <- 100*CO_req_send_age$Count/CO_req_send_age[7,2]

# Florida

FL_req_send_party <- party_shell

FL_req_send_party[1,2] <- state_stats[10,10] 
FL_req_send_party[2,2] <- state_stats[10,11] 
FL_req_send_party[3,2] <- state_stats[10,12] 
FL_req_send_party[4,2] <- state_stats[10,13] 
FL_req_send_party[5,2] <- state_stats[10,5] 

FL_req_send_party$Percent <- 100*FL_req_send_party$Count/FL_req_send_party[5,2]

# Illinois

IL_req_send_tot <- data.frame(Total=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)

IL_req_send_tot[1,1] <- "TOTAL"
IL_req_send_tot[1,2] <- sum(state_stats[14,5]) 

IL_req_send_tot$Percent <- 100.0

# North Carolina

NC_req_send_party <- party_shell

NC_req_send_party[1,2] <- state_stats[34,10] 
NC_req_send_party[2,2] <- state_stats[34,11] 
NC_req_send_party[3,2] <- state_stats[34,12] 
NC_req_send_party[4,2] <- state_stats[34,13] 
NC_req_send_party[5,2] <- state_stats[34,5] 

NC_req_send_party$Percent <- 100*NC_req_send_party$Count/NC_req_send_party[5,2]

NC_req_send_race <- race_shell

NC_req_send_race[1,2] <- sum(state_stats[34,14])
NC_req_send_race[2,2] <- sum(state_stats[34,15]) 
NC_req_send_race[3,2] <- sum(state_stats[34,16]) 
NC_req_send_race[4,2] <- sum(state_stats[34,17]) 
NC_req_send_race[5,2] <- sum(state_stats[34,18]) 
NC_req_send_race[6,2] <- sum(state_stats[34,19]) 
NC_req_send_race[7,2] <- sum(state_stats[34,5]) 

NC_req_send_race$Percent <- 100 * NC_req_send_race$Count/NC_req_send_race[7,2]

NC_req_send_gender <- gender_shell

NC_req_send_gender[1,2] <- sum(state_stats[34,21])
NC_req_send_gender[2,2] <- sum(state_stats[34,20]) 
NC_req_send_gender[3,2] <- sum(state_stats[34,22]) 
NC_req_send_gender[4,2] <- sum(state_stats[34,5]) 

NC_req_send_gender$Percent <- 100 * NC_req_send_gender$Count/NC_req_send_gender[4,2]

# Ohio

OH_req_send_tot <- data.frame(Total=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)

OH_req_send_tot[1,1] <- "TOTAL"
OH_req_send_tot[1,2] <- sum(state_stats[36,5]) 

OH_req_send_tot$Percent <- 100.0

OH_req_send_age <- age_shell

OH_req_send_age[1,2] <- state_stats[36,23] 
OH_req_send_age[2,2] <- state_stats[36,24] 
OH_req_send_age[3,2] <- state_stats[36,25] 
OH_req_send_age[4,2] <- state_stats[36,26] 
OH_req_send_age[5,2] <- state_stats[36,27] 
OH_req_send_age[6,2] <- state_stats[36,28] 
OH_req_send_age[7,2] <- state_stats[36,5] 

OH_req_send_age$Percent <- 100*OH_req_send_age$Count/OH_req_send_age[7,2]


# Pennsylvania

PA_req_send_party <- party_shell

PA_req_send_party[1,2] <- state_stats[39,10] 
PA_req_send_party[2,2] <- state_stats[39,11] 
PA_req_send_party[3,2] <- state_stats[39,12] 
PA_req_send_party[4,2] <- state_stats[39,13] 
PA_req_send_party[5,2] <- state_stats[39,5] 

PA_req_send_party$Percent <- 100*PA_req_send_party$Count/PA_req_send_party[5,2]

PA_req_send_age <- age_shell

PA_req_send_age[1,2] <- state_stats[39,23] 
PA_req_send_age[2,2] <- state_stats[39,24] 
PA_req_send_age[3,2] <- state_stats[39,25] 
PA_req_send_age[4,2] <- state_stats[39,26] 
PA_req_send_age[5,2] <- state_stats[39,27] 
PA_req_send_age[6,2] <- state_stats[39,28] 
PA_req_send_age[7,2] <- state_stats[39,5] 

PA_req_send_age$Percent <- 100*PA_req_send_age$Count/PA_req_send_age[7,2]

# South Carolina

SC_req_send_race <- race_shell

SC_req_send_race[1,2] <- sum(state_stats[41,14])
SC_req_send_race[2,2] <- sum(state_stats[41,15]) 
SC_req_send_race[3,2] <- sum(state_stats[41,16]) 
SC_req_send_race[4,2] <- sum(state_stats[41,17]) 
SC_req_send_race[5,2] <- sum(state_stats[41,18]) 
SC_req_send_race[6,2] <- sum(state_stats[41,19]) 
SC_req_send_race[7,2] <- sum(state_stats[41,5]) 

SC_req_send_race$Percent <- 100 * SC_req_send_race$Count/SC_req_send_race[7,2]
```
Michael McDonald  
Professor, University of Florida  
Web: [United States Elections Project](http://www.electproject.org/)  
Twitter: [\@ElectProject](https://twitter.com/ElectProject)

Last updated: `r format(as.POSIXct(Sys.time(), format = "%y%m%d %H:%M"), "%m/%d/%Y %I:%M %p")` Eastern Time

## **National Statistics**

There are a total of **`r format(as.numeric(sum(state_stats$mail_sent_req_2020)), big.mark =",")`** ballot requests in the Reporting states of CO, FL, IL, NC, OH (partial), PA, and SC.

Scroll down for more state statistics and links to these state data, where publicly available.

### Mail Ballot Requests by Party Registration

Reporting states with party registration data: CO, FL, NC, PA  

``` {r echo = FALSE}

options(knitr.kable.NA = '')

kable(US_req_send_party, valign = 't', format.args = list(big.mark = ",", scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
Registered Democrats have a **`r format(as.numeric(US_req_send_party[1,2]-US_req_send_party[2,2]), big.mark =",")`** ballot request lead over registered Republicans.

*Note:* Party registration statistics are for states that have party registration. These statistics are not actual votes. By federal law, election officials do not begin counting ballots until Election Day, although they may start the process of preparing ballots for counting in advance. 

### Mail Ballot Requests by Race and Ethnicity

Reporting states with race or ethnicity data: NC, SC   

``` {r echo = FALSE}

options(knitr.kable.NA = '')

kable(US_req_send_race, valign = 't', format.args = list(big.mark = ",", scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

*Note:* Race and ethnicity statistics are for states that ask registered voters to provide their race and ethnicity. 

### Mail Ballot Requests by Age

Reporting states with age data: CO, OH (partial), PA   

``` {r echo = FALSE}
kable(US_req_send_age, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```


### Mail Ballot Requests by Gender

Reporting states with gender data: CO, NC   

``` {r echo = FALSE}
kable(US_req_send_gender, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

***

## **Colorado**

Last Report: `r state_stats[6,9]`  
Source:  `r state_stats[6,2]`

### Discussion

Colorado is an all-mail ballot state. Ballot requests reported here are counts of the state's registered voters

### Mail Ballot Requests by Party Registration

``` {r echo = FALSE}
kable(CO_req_send_party,format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

Colorado registered Democrats have a **`r format(as.numeric(CO_req_send_party[1,2]-CO_req_send_party[2,2]), big.mark =",")`** ballot request lead over registered Republicans.

### Mail Ballot Requests by Age

``` {r echo = FALSE}
kable(CO_req_send_age, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

### Mail Ballot Requests by Gender

``` {r echo = FALSE}
kable(CO_req_send_gender, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

***

## **Florida**

Last Report: `r state_stats[10,9]`  
Source:  `r state_stats[10,2]`

### Discussion

Florida's ballot requests are good through two general elections. All August primary ballot requests are good for November, unless a voter notifies election officials they wish to cancel their request.

To calculate the number of requested ballots from the Florida Division of Elections website, add the rows *Vote-by-Mail Provided (Not Yet Returned)* and *Voted Vote-by-Mail*.

### Mail Ballot Requests by Party Registration

``` {r echo = FALSE}
kable(FL_req_send_party,format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

Florida registered Democrats have a **`r format(as.numeric(FL_req_send_party[1,2]-FL_req_send_party[2,2]), big.mark =",")`** ballot request lead over registered Republicans.

***

## **Illinois**

Last Report: `r state_stats[14,9]`  
Source:  `r state_stats[14,2]`

### Discussion

Illinois provides only early ballot counts by county.

### Mail Ballot Requests

``` {r echo = FALSE}
kable(IL_req_send_tot, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

***

## **North Carolina**

Last Report: `r state_stats[34,9]`  
Source:  `r state_stats[34,2]`

### Discussion

North Carolina is the first state to begin general election voting. Election officials start sending ballots to voters on September 4.

### Mail Ballot Requests by Party Registration

``` {r echo = FALSE}
kable(NC_req_send_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

North Carolina registered Democrats have a **`r format(as.numeric(NC_req_send_party[1,2]-NC_req_send_party[2,2]), big.mark =",")`** ballot request lead over registered Republicans.

### Mail Ballot Requests by Race and Ethnicity

``` {r echo = FALSE}
kable(NC_req_send_race, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

### Mail Ballot Requests by Gender

``` {r echo = FALSE}
kable(NC_req_send_gender, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

***

## **Ohio**

Last Report: `r state_stats[36,9]`  
Source:  `r state_stats[36,2]`

### Discussion

Data presented here are scraped from 62 of Ohio's 88 counties that post absentee ballot reports online. These 62 reporting counties include the largest counties of Cuyahoga, Franklin, and Hamilton.  

These data were last scraped on the Last Report date, but counties may have last posted their data on an earlier date.

You may see reports of Ohio mail ballot activity by "Party". I will not post party statistics because they are misleading. Ohio's party affiliation is unlike party registration in other states. In Ohio, "Party" is a record of the last party primary a voter participated in, in the current and past last two calendar years (see: [Ohio Code 3513.19(3)](https://codes.ohio.gov/orc/3513.19)). In Ohio's April 28, 2020 presidential primary, 124,628 more voters participated in the Democratic primary than the Republican primary, primarily because President Trump had only token opposition. As a consequence, **Ohio's "Party" is currently artificially tilted towards the Democrats.** Leveraging this quirk, both political parties and their allies have abused these party statistics over the years to make disingenuous claims about party enthusiasm in Ohio.

### Mail Ballot Requests

``` {r echo = FALSE}
kable(OH_req_send_tot, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

### Mail Ballot Requests by Age

``` {r echo = FALSE}
kable(OH_req_send_age, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

***

## **Pennsylvania**

Last Report: `r state_stats[39,9]`  
Source:  `r state_stats[39,2]`

### Discussion

Pennsylvania ballot requests are good through the end of a calendar year unless a voter otherwise cancels it. I analyze the ballot requests for the state's closed primary. The number of ballot requests from all voters, especially minor party and no party affiliated voters, is likely larger than reported here.

### Mail Ballot Requests by Party Registration

``` {r echo = FALSE}
kable(PA_req_send_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

Pennsylvania registered Democrats have a **`r format(as.numeric(PA_req_send_party[1,2]-PA_req_send_party[2,2]), big.mark =",")`** ballot request lead over registered Republicans.

### Mail Ballot Requests by Age

``` {r echo = FALSE}
kable(PA_req_send_age, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```


***

## **South Carolina**

Last Report: `r state_stats[41,9]`  
Source:  `r state_stats[41,2]`

### Discussion

South Carolina does not have party registration, but does provides ballot counts by race.

### Mail Ballot Requests by Race and Ethnicity

``` {r echo = FALSE}
kable(SC_req_send_race, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
