---
title: "Early Vote Analysis for Sunday, Oct. 25"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(DT)
library(highcharter)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

party_shell_returned <- data.frame(Party=character(),
                                  Count=integer(),
                                  Frequency=double(),
                                  Count2=integer(),
                                  Rate=integer(),
                                  stringsAsFactors=FALSE)
party_shell_returned[1,1] <- "Democrats"
party_shell_returned[2,1] <- "Republicans"
party_shell_returned[3,1] <- "Minor"
party_shell_returned[4,1] <- "No Party Affiliation"
party_shell_returned[5,1] <- "TOTAL"


US_accept_party <- party_shell_returned

US_accept_party[1,2] <- sum(state_stats$mail_accept_2020_dem) 
US_accept_party[2,2] <- sum(state_stats$mail_accept_2020_rep) 
US_accept_party[3,2] <- sum(state_stats$mail_accept_2020_minor)
US_accept_party[4,2] <- sum(state_stats$mail_accept_2020_npa) 
US_accept_party[5,2] <- sum(state_stats$mail_accept_2020_dem)+sum(state_stats$mail_accept_2020_rep)+sum(state_stats$mail_accept_2020_minor)+sum(state_stats$mail_accept_2020_npa)

US_accept_party[1,4] <- sum(state_stats$mail_sent_req_2020_dem) 
US_accept_party[2,4] <- sum(state_stats$mail_sent_req_2020_rep) 
US_accept_party[3,4] <- sum(state_stats$mail_sent_req_2020_minor)
US_accept_party[4,4] <- sum(state_stats$mail_sent_req_2020_npa) 
US_accept_party[5,4] <- sum(state_stats$mail_sent_req_2020_dem)+sum(state_stats$mail_sent_req_2020_rep)+sum(state_stats$mail_sent_req_2020_minor)+sum(state_stats$mail_sent_req_2020_npa) 

US_accept_party$Frequency <- 100*US_accept_party$Count/US_accept_party[5,2]
US_accept_party$Rate <- 100*US_accept_party$Count/US_accept_party$Count2

colnames(US_accept_party) <- c("Party", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")

US_out_party <- party_shell_returned

US_out_party[1,2] <- sum(state_stats$mail_sent_req_2020_dem) - sum(state_stats$mail_accept_2020_dem) 
US_out_party[2,2] <- sum(state_stats$mail_sent_req_2020_rep) - sum(state_stats$mail_accept_2020_rep) 
US_out_party[3,2] <- sum(state_stats$mail_sent_req_2020_minor) - sum(state_stats$mail_accept_2020_minor)
US_out_party[4,2] <- sum(state_stats$mail_sent_req_2020_npa) - sum(state_stats$mail_accept_2020_npa) 
US_out_party[5,2] <- (sum(state_stats$mail_sent_req_2020_dem)+sum(state_stats$mail_sent_req_2020_rep)+sum(state_stats$mail_sent_req_2020_minor)+sum(state_stats$mail_sent_req_2020_npa)) - (sum(state_stats$mail_accept_2020_dem)+sum(state_stats$mail_accept_2020_rep)+sum(state_stats$mail_accept_2020_minor)+sum(state_stats$mail_accept_2020_npa))

US_out_party[1,5] <- sum(state_stats$mail_sent_req_2020_dem) 
US_out_party[2,5] <- sum(state_stats$mail_sent_req_2020_rep) 
US_out_party[3,5] <- sum(state_stats$mail_sent_req_2020_minor)
US_out_party[4,5] <- sum(state_stats$mail_sent_req_2020_npa) 
US_out_party[5,5] <- sum(state_stats$mail_sent_req_2020_dem)+sum(state_stats$mail_sent_req_2020_rep)+sum(state_stats$mail_sent_req_2020_minor)+sum(state_stats$mail_sent_req_2020_npa) 

US_out_party[1,4] <- sum(state_stats$mail_accept_2020_dem) 
US_out_party[2,4] <- sum(state_stats$mail_accept_2020_rep) 
US_out_party[3,4] <- sum(state_stats$mail_accept_2020_minor)
US_out_party[4,4] <- sum(state_stats$mail_accept_2020_npa) 
US_out_party[5,4] <- sum(state_stats$mail_accept_2020_dem)+sum(state_stats$mail_accept_2020_rep)+sum(state_stats$mail_accept_2020_minor)+sum(state_stats$mail_accept_2020_npa)

US_out_party$Frequency <- 100*US_out_party$Count/US_out_party[5,2]

colnames(US_out_party) <- c("Party", "Outstanding Ballots", "Freq. Distribution", "Returned Ballots", "Requested Ballots")

# Voted

US_voted_party <- party_shell

US_voted_party[1,2] <- sum(state_stats$mail_accept_2020_dem) + sum(state_stats$inperson_2020_dem)
US_voted_party[2,2] <- sum(state_stats$mail_accept_2020_rep) + sum(state_stats$inperson_2020_rep)
US_voted_party[3,2] <- sum(state_stats$mail_accept_2020_minor) + sum(state_stats$inperson_2020_minor)
US_voted_party[4,2] <- sum(state_stats$mail_accept_2020_npa) + sum(state_stats$inperson_2020_npa)
US_voted_party[5,2] <- sum(state_stats$mail_accept_2020_dem)+sum(state_stats$mail_accept_2020_rep)+sum(state_stats$mail_accept_2020_minor)+sum(state_stats$mail_accept_2020_npa) + sum(state_stats$inperson_2020_dem)+sum(state_stats$inperson_2020_rep)+sum(state_stats$inperson_2020_minor)+sum(state_stats$inperson_2020_npa)

US_voted_party$Percent <- 100*US_voted_party$Count/US_voted_party[5,2]

# Inperson

US_inperson_party <- party_shell

US_inperson_party[1,2] <- sum(state_stats$inperson_2020_dem) 
US_inperson_party[2,2] <- sum(state_stats$inperson_2020_rep) 
US_inperson_party[3,2] <- sum(state_stats$inperson_2020_minor)
US_inperson_party[4,2] <- sum(state_stats$inperson_2020_npa) 
US_inperson_party[5,2] <- sum(state_stats$inperson_2020_dem)+sum(state_stats$inperson_2020_rep)+sum(state_stats$inperson_2020_minor)+sum(state_stats$inperson_2020_npa)

US_inperson_party$Percent <- 100*US_inperson_party$Count/US_inperson_party[5,2]

NM_inperson_party <- party_shell

NM_inperson_party[1,2] <- state_stats[32,77] 
NM_inperson_party[2,2] <- state_stats[32,78] 
NM_inperson_party[3,2] <- state_stats[32,79] 
NM_inperson_party[4,2] <- state_stats[32,80] 
NM_inperson_party[5,2] <- state_stats[32,7] 

NM_inperson_party$Percent <- 100*NM_inperson_party$Count/NM_inperson_party[5,2]

NM_accept_party <- party_shell_returned

NM_accept_party[1,2] <- state_stats[32,29] 
NM_accept_party[2,2] <- state_stats[32,30] 
NM_accept_party[3,2] <- state_stats[32,31] 
NM_accept_party[4,2] <- state_stats[32,32] 
NM_accept_party[5,2] <- state_stats[32,6] 

NM_accept_party[1,4] <- state_stats[32,10] 
NM_accept_party[2,4] <- state_stats[32,11] 
NM_accept_party[3,4] <- state_stats[32,12] 
NM_accept_party[4,4] <- state_stats[32,13] 
NM_accept_party[5,4] <- state_stats[32,5] 

NM_accept_party$Frequency <- 100*NM_accept_party$Count/NM_accept_party[5,2]
NM_accept_party$Rate <- 100*NM_accept_party$Count/NM_accept_party$Count2

colnames(NM_accept_party) <- c("Party", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")

NC_inperson_party <- party_shell

NC_inperson_party[1,2] <- state_stats[34,77] 
NC_inperson_party[2,2] <- state_stats[34,78] 
NC_inperson_party[3,2] <- state_stats[34,79] 
NC_inperson_party[4,2] <- state_stats[34,80] 
NC_inperson_party[5,2] <- state_stats[34,7] 

NC_inperson_party$Percent <- 100*NC_inperson_party$Count/NC_inperson_party[5,2]

NC_accept_party <- party_shell_returned

NC_accept_party[1,2] <- state_stats[34,29] 
NC_accept_party[2,2] <- state_stats[34,30] 
NC_accept_party[3,2] <- state_stats[34,31] 
NC_accept_party[4,2] <- state_stats[34,32] 
NC_accept_party[5,2] <- state_stats[34,6] 

NC_accept_party[1,4] <- state_stats[34,10] 
NC_accept_party[2,4] <- state_stats[34,11] 
NC_accept_party[3,4] <- state_stats[34,12] 
NC_accept_party[4,4] <- state_stats[34,13] 
NC_accept_party[5,4] <- state_stats[34,5] 

NC_accept_party$Frequency <- 100*NC_accept_party$Count/NC_accept_party[5,2]
NC_accept_party$Rate <- 100*NC_accept_party$Count/NC_accept_party$Count2

colnames(NC_accept_party) <- c("Party", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")

LA_inperson_party <- party_shell

LA_inperson_party[1,2] <- state_stats[19,77] 
LA_inperson_party[2,2] <- state_stats[19,78] 
LA_inperson_party[3,2] <- state_stats[19,79] 
LA_inperson_party[4,2] <- state_stats[19,80] 
LA_inperson_party[5,2] <- state_stats[19,7] 

LA_inperson_party$Percent <- 100*LA_inperson_party$Count/LA_inperson_party[5,2]

FL_inperson_party <- party_shell

FL_inperson_party[1,2] <- state_stats[10,77] 
FL_inperson_party[2,2] <- state_stats[10,78] 
FL_inperson_party[3,2] <- state_stats[10,79] 
FL_inperson_party[4,2] <- state_stats[10,80] 
FL_inperson_party[5,2] <- state_stats[10,7] 

FL_inperson_party$Percent <- 100*FL_inperson_party$Count/FL_inperson_party[5,2]


FL_voted_party <- party_shell

FL_voted_party[1,2] <- state_stats[10,77] + state_stats[10,29] 
FL_voted_party[2,2] <- state_stats[10,78] + state_stats[10,30] 
FL_voted_party[3,2] <- state_stats[10,79] + state_stats[10,31]
FL_voted_party[4,2] <- state_stats[10,80] + state_stats[10,32] 
FL_voted_party[5,2] <- state_stats[10,7] + state_stats[10,6]

FL_voted_party$Percent <- 100*FL_voted_party$Count/FL_voted_party[5,2]

FL_return_party <- party_shell_returned

FL_return_party[1,2] <- state_stats[10,29] 
FL_return_party[2,2] <- state_stats[10,30] 
FL_return_party[3,2] <- state_stats[10,31] 
FL_return_party[4,2] <- state_stats[10,32] 
FL_return_party[5,2] <- state_stats[10,6] 

FL_return_party[1,4] <- state_stats[10,10] 
FL_return_party[2,4] <- state_stats[10,11] 
FL_return_party[3,4] <- state_stats[10,12] 
FL_return_party[4,4] <- state_stats[10,13] 
FL_return_party[5,4] <- state_stats[10,5] 

FL_return_party$Frequency <- 100*FL_return_party$Count/FL_return_party[5,2]
FL_return_party$Rate <- 100*FL_return_party$Count/FL_return_party$Count2

colnames(FL_return_party) <- c("Party", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")


```

## **Early Voting Continues at a Record Pace**

The numbers are stunning. Almost 60 million people have already voted as of Sunday, October 25. That is more than the [47.2 million 2016 general election pre-election votes](http://www.electproject.org/early_2016), as I stopped collecting statistics on Election Day in 2016. It is more than the [58 million in-person and mail ballots cast in 2016](https://apnews.com/article/election-2020-donald-trump-politics-florida-elections-509ad83f6d40e08fb715da44548f62e0), which includes mail ballots that were dropped off on Election Day or post marked by Election Day and received by election offices afterwards (according to state laws -- be sure of your state's laws if you plan to vote this way).

Nationally, the 2020 early vote is greater than 43% of all votes cast in the 2016 election. The pace of some states' early voting is such that with almost certainty states will begin surpassing their total 2016 total vote this week.

| State         | % 2016 Total Vote |
| ------------- |-------------:|
| Texas         | 80% |
| Montana       | 70% |
| North Carolina | 65% |
| Tennessee | 65% |
| Georgia | 65% |
| New Mexico | 60% |
| Hawaii (with only a Honolulu County report!) | 60% |
| Washington | 60% |
| Florida | 60% |

Washington is illuminating about what is happening. The state ran all-mail elections in 2016 and 2020 under the same rules. The only change is a change in voters' behavior. So far, the [state reports](https://www.sos.wa.gov/elections/research/ballot-return-statistics.aspx) 2,018,495 mail ballots have been returned. At this same number of days prior to the election in 2016, 763,934 ballots were returned. Washington's mail ballot return rates are nearly three times what they were in 2016.

Nationwide, voters will not only be sent an unprecedented number of at least 87 million mail ballots, but they are returning them sooner than in past elections. In all nearly 40 million mail ballots have been returned so far, a return rate of nearly 46%.

**This is good news!** There were many concerns about election officials' ability to conduct an election during a pandemic. Not only are people voting, but they are voting over a longer period of time, thereby spreading out the workload of election officials. Yes, there have been problems, and in many places lines are intolerably long. But, people are voting and there are more opportunities for them to do so by Election Day. Americans' resilience and support for their democracy is very heartening in these trying times.

## **Democrats and Republicans Have Different Concerns Moving Forward**

The election is not over yet, not by a long shot. There is still a little more than a week before the election, and Election Day itself to go.

Democrats and Republicans will commonly fret about their voters, but their concerns are in different places.

### **Democrats**

Democrats enjoy a huge advantage in mail ballot requests, and mail ballot return rates. These national numbers for the states with party registration are reflected in every party registration state; they are not just an artifact of heavily Democratic large states like California conducting all-mail ballot elections.

#### Returned Mail Ballots by Party Registration
``` {r echo = FALSE}

options(knitr.kable.NA = '')

kable(US_accept_party, valign = 't', format.args = list(big.mark = ",", scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

But let's turnout the mail ballot phenomenon on its head. Voters still haven't returned a little more than 47 million mail ballots to election officials. 

Again, let's take a look at the party registation states.

#### Outstanding Mail Ballots by Party Registration
Reporting states with party registration data: CA, CO, CT, FL, IA, KS, KY, MD, ME, NC, NE, NJ, NM, NV, OK, OR, PA, SD  
``` {r echo = FALSE}

options(knitr.kable.NA = '')

kable(US_out_party, valign = 't', format.args = list(big.mark = ",", scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

Because registered Democrats are voting in such large numbers by mail, even though their ballot return rate is higher than Republicans, they have more than 4 million mail ballots outstanding. This is just the states with party registration; we might surmise the total differential number of Democrats with outstanding mail ballots is higher.

### **Republicans**

Republicans need to vote in-person to make up ground on the Democratic mail voters, either early or on Election Day. There is still some play left in the in-person early vote, but time is starting to run short such that Republicans will need to rely heavily on Election Day vote, which has traditionally been a strong day of voting for Republicans in recent elections.

Again, we turn to the reporting states with party registration. Republicans have a lead in the in-person early vote, but it is not as large as the Democrats' lead in the mail ballots, nor is it the margin as large.

**In-Person Votes by Party Registration**  
Reporting states with party registration data: CO, FL, KS, KY, LA, MD, NC, NM, NV, SD  
``` {r echo = FALSE}

options(knitr.kable.NA = '')

kable(US_inperson_party, valign = 't', format.args = list(big.mark = ",", scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

Still, there could be a bias from the selected states. Louisiana did not expand mail balloting, so more voters are casting ballots in-person, which is also happening in Texas. Texas -- which like Louisiana did not expand mail balloting options -- probably looks similar to Louisiana. (Louisiana also doesn't differentiate between in-person and mail ballots in its reporting of party registration statistics; mail ballot are only about a quarter of the reported early votes.)

**Louisiana In-Person Votes by Party Registration**  
``` {r echo = FALSE}

options(knitr.kable.NA = '')

kable(LA_inperson_party, valign = 't', format.args = list(big.mark = ",", scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

North Carolina is another interesting case, where Republicans are doing better among in-person early voters, but Democrats still have a lead. I'll point you to the [North Carolina dashboard](https://electproject.github.io/Early-Vote-2020G/NC.html) for supporting statistics. An election law that is different in North Carolina is same day registration during the in-person early voting period. It may be that new voters are adding to the Democratic numbers of in-person early voters.

In the critical battleground state of Florida, Republicans are making up some ground on Democrats through the in-person early vote. 

**Florida Mail Ballots Returned by Party Registration**
``` {r echo = FALSE}
kable(FL_return_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

**Florida In-Person Votes by Party Registration**  
``` {r echo = FALSE}

options(knitr.kable.NA = '')

kable(FL_inperson_party, valign = 't', format.args = list(big.mark = ",", scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

The net sum is still a sizable Democratic lead, but not as large as it was before in-person early voting started. Similar patterns are evident in the remaining party registration states reporting in-person early voting statistics.

**Florida Total Voted by Party Registration**  
``` {r echo = FALSE}

options(knitr.kable.NA = '')

kable(FL_voted_party, valign = 't', format.args = list(big.mark = ",", scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

Florida election officials must provide early voting through Friday, and have the discretion to extend the days until Sunday. There are a lot of Democratic mail ballots still outstanding. Will Democrats return them by Election Day, as is required for all but military and overseas civilian voters? How much will Republicans make up ground with in-person early voting? Will African-Americans show up to vote in-person early during Sunday's Souls-to-Polls events, offsetting some Republican in-person gains? What will happen on Election Day? 

Answers to these questions should come into sharper focus by my last update next week, where I plan to have more precise turnout forecasts and a couple of projections for all-mail states like Colorado, Oregon, and Nevada. And before you think these states are uninteresting because they aren't all battlegrounds, they may provide us with a sense of the direction and magnitude of the national swing towards Biden or Trump.