---
title: "Florida Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

# Florida

FL_req_send_party <- party_shell

FL_req_send_party[1,2] <- state_stats[10,10] 
FL_req_send_party[2,2] <- state_stats[10,11] 
FL_req_send_party[3,2] <- state_stats[10,12] 
FL_req_send_party[4,2] <- state_stats[10,13] 
FL_req_send_party[5,2] <- state_stats[10,5] 

FL_req_send_party$Percent <- 100*FL_req_send_party$Count/FL_req_send_party[5,2]
```
Last Report: `r state_stats[10,9]`  
Source:  `r state_stats[10,2]`

### Discussion

Florida's ballot requests are good through two general elections. All August primary ballot requests are good for November, unless a voter notifies election officials they wish to cancel their request.

To calculate the number of requested ballots from the Florida Division of Elections website, add the rows *Vote-by-Mail Provided (Not Yet Returned)* and *Voted Vote-by-Mail*.

### Mail Ballot Requests by Party Registration

``` {r echo = FALSE}
kable(FL_req_send_party,format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

Florida registered Democrats have a **`r format(as.numeric(FL_req_send_party[1,2]-FL_req_send_party[2,2]), big.mark =",")`** ballot request lead over registered Republicans.
