---
title: "Georgia Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(DT)
library(highcharter)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")
GA_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote_GA.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"

race_shell_returned <- data.frame(Race=character(),
                                  Count=integer(),
                                  Frequency=double(),
                                  Count2=integer(),
                                  Rate=integer(),
                                  stringsAsFactors=FALSE)
race_shell_returned[1,1] <- "Non-Hispanic White"
race_shell_returned[2,1] <- "Non-Hispanic Black"
race_shell_returned[3,1] <- "Hispanic"
race_shell_returned[4,1] <- "Non-Hispanic Asian American"
race_shell_returned[5,1] <- "Non-Hispanic Native American"
race_shell_returned[6,1] <- "Other/Multiple/Unknown"
race_shell_returned[7,1] <- "TOTAL"

GA_age_shell_returned <- data.frame(Race=character(),
                                  Count=integer(),
                                  Frequency=double(),
                                  Count2=integer(),
                                  Rate=integer(),
                                  stringsAsFactors=FALSE)
GA_age_shell_returned[1,1] <- "18 to 24"
GA_age_shell_returned[2,1] <- "25 to 34"
GA_age_shell_returned[3,1] <- "35 to 44"
GA_age_shell_returned[4,1] <- "45 and 55"
GA_age_shell_returned[5,1] <- "56 and 65"
GA_age_shell_returned[6,1] <- "66 and up"
GA_age_shell_returned[7,1] <- "Age Unknown"
GA_age_shell_returned[8,1] <- "TOTAL"

gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

# Georiga

GA_accept_race <- race_shell_returned

GA_accept_race[1,2] <- sum(state_stats[11,33])
GA_accept_race[2,2] <- sum(state_stats[11,34]) 
GA_accept_race[3,2] <- sum(state_stats[11,37]) 
GA_accept_race[4,2] <- sum(state_stats[11,35]) 
GA_accept_race[5,2] <- sum(state_stats[11,36]) 
GA_accept_race[6,2] <- sum(state_stats[11,38]) 
GA_accept_race[7,2] <- sum(state_stats[11,6]) 

GA_accept_race[1,4] <- sum(state_stats[11,14])
GA_accept_race[2,4] <- sum(state_stats[11,15]) 
GA_accept_race[3,4] <- sum(state_stats[11,16]) 
GA_accept_race[4,4] <- sum(state_stats[11,17]) 
GA_accept_race[5,4] <- sum(state_stats[11,18]) 
GA_accept_race[6,4] <- sum(state_stats[11,19]) 
GA_accept_race[7,4] <- sum(state_stats[11,5]) 

GA_accept_race$Frequency <- 100 * GA_accept_race$Count/GA_accept_race[7,2]
GA_accept_race$Rate <- 100*GA_accept_race$Count/GA_accept_race$Count2

colnames(GA_accept_race) <- c("Race/Ethnicity", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")

GA_accept_age <- GA_age_shell_returned

GA_accept_age[1,2] <- sum(state_stats[11,42])
GA_accept_age[2,2] <- sum(state_stats[11,43]) 
GA_accept_age[3,2] <- sum(state_stats[11,44]) 
GA_accept_age[4,2] <- sum(state_stats[11,45]) 
GA_accept_age[5,2] <- sum(state_stats[11,46])
GA_accept_age[6,2] <- sum(state_stats[11,47])
GA_accept_age[7,2] <- sum(state_stats[11,75])
GA_accept_age[8,2] <- sum(state_stats[11,6])

GA_accept_age[1,4] <- sum(state_stats[11,23])
GA_accept_age[2,4] <- sum(state_stats[11,24]) 
GA_accept_age[3,4] <- sum(state_stats[11,25]) 
GA_accept_age[4,4] <- sum(state_stats[11,26]) 
GA_accept_age[5,4] <- sum(state_stats[11,27])
GA_accept_age[6,4] <- sum(state_stats[11,28])
GA_accept_age[7,4] <- sum(state_stats[11,74])
GA_accept_age[8,4] <- sum(state_stats[11,5])

GA_accept_age$Frequency <- 100 * GA_accept_age$Count/GA_accept_age[5,2]
GA_accept_age$Rate <- 100*GA_accept_age$Count/GA_accept_age$Count2

colnames(GA_accept_age) <- c("Age", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")

GA_req_send_tot <- data.frame(Total=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)

GA_req_send_tot[1,1] <- "TOTAL"
GA_req_send_tot[1,2] <- sum(state_stats[11,5]) 

GA_req_send_tot$Percent <- 100.0


GA_stats_requests <- select(GA_stats, County, Reg.Voters, Mail.Req.Tot, Pct.Request)
GA_stats_returns <- select(GA_stats, County, Mail.Req.Tot, Mail.Accept.Tot, Pct.Accept)
```
Last Report: `r state_stats[11,9]`  
Source:  `r state_stats[11,2]`

## {.tabset}

### Mail Ballots Returned and Accepted
Mail Ballots Returned and Accepted: **`r format(as.numeric(sum(GA_stats$Mail.Accept.Tot)), big.mark =",")`**  
Return Rate: **`r paste(round(100*sum(GA_stats$Mail.Accept.Tot)/sum(GA_stats$Mail.Req.Tot), digits = 1),"%", sep = "")`**

``` {r echo = FALSE}
GA_map_data <- GA_stats

GA_map_data <- mutate(GA_map_data, percent = round(100*(Mail.Accept.Tot/Mail.Req.Tot), digits = 1))

# GA_map_data <- rename(GA_map_data, fips = FIPS_CODE)
GA_map_data <- mutate(GA_map_data, fips = as.character(FIPS_CODE))

mapfile <- download_map_data("countries/us/us-ga-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

GA_map_data <- left_join(GA_map_data, mapdata, by = "fips")

GA_map_data <- arrange(GA_map_data, row)

hcmap(map = "countries/us/us-ga-all", data = GA_map_data,
      value = "percent", name = "Percent Accepted", joinBy = "fips") 
```

``` {r echo = FALSE}
datatable(GA_stats_returns, colnames = c("County", "Mail Ballots Requested", "Mail Ballots Returned", "Percent Returned"), rownames = F) %>%
  formatPercentage('Pct.Accept', 1) %>%
  formatRound(c('Mail.Req.Tot', 'Mail.Accept.Tot'), 0, mark = ",")
```

#### Mail Ballots Returned and Accepted by Race and Ethnicity

``` {r echo = FALSE}
kable(GA_accept_race, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

#### Mail Ballots Returned and Accepted by Age

``` {r echo = FALSE}
kable(GA_accept_age, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```


### Requested

Ballots Requested: **`r format(as.numeric(sum(GA_stats$Mail.Req.Tot)), big.mark =",")`**  
Request Rate: **`r paste(round(100*sum(GA_stats$Mail.Req.Tot)/sum(GA_stats$Reg.Voters), digits = 1),"%", sep = "")`**

``` {r echo = FALSE}
GA_map_data <- GA_stats

GA_map_data <- mutate(GA_map_data, percent = round(100*(Mail.Req.Tot/Reg.Voters), digits = 1))

# GA_map_data <- rename(GA_map_data, fips = FIPS_CODE)
GA_map_data <- mutate(GA_map_data, fips = as.character(FIPS_CODE))

mapfile <- download_map_data("countries/us/us-ga-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

GA_map_data <- left_join(GA_map_data, mapdata, by = "fips")

GA_map_data <- arrange(GA_map_data, row)

hcmap(map = "countries/us/us-ga-all", data = GA_map_data,
      value = "percent", name = "Percent Requested", joinBy = "fips") 
```

``` {r echo = FALSE}
datatable(GA_stats_requests, colnames = c("County", "Registered Voters", "Mail Ballots Requested", "Percent Requested"), rownames = F) %>%
  formatPercentage('Pct.Request', 1) %>%
  formatRound(c('Reg.Voters','Mail.Req.Tot'), 0, mark = ",")
```
