---
title: "Iowa Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(DT)
library(highcharter)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")
IA_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote_IA.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

party_shell_returned <- data.frame(Party=character(),
                                  Count=integer(),
                                  Frequency=double(),
                                  Count2=integer(),
                                  Rate=integer(),
                                  stringsAsFactors=FALSE)
party_shell_returned[1,1] <- "Democrats"
party_shell_returned[2,1] <- "Republicans"
party_shell_returned[3,1] <- "Minor"
party_shell_returned[4,1] <- "No Party Affiliation"
party_shell_returned[5,1] <- "TOTAL"

# Iowa

IA_req_send_party <- party_shell

IA_req_send_party[1,2] <- state_stats[16,10] 
IA_req_send_party[2,2] <- state_stats[16,11] 
IA_req_send_party[3,2] <- state_stats[16,12] 
IA_req_send_party[4,2] <- state_stats[16,13] 
IA_req_send_party[5,2] <- state_stats[16,5] 

IA_req_send_party$Percent <- 100*IA_req_send_party$Count/IA_req_send_party[5,2]

IA_accept_party <- party_shell_returned

IA_accept_party[1,2] <- state_stats[16,29] 
IA_accept_party[2,2] <- state_stats[16,30] 
IA_accept_party[3,2] <- state_stats[16,31] 
IA_accept_party[4,2] <- state_stats[16,32] 
IA_accept_party[5,2] <- state_stats[16,6] 

IA_accept_party[1,4] <- state_stats[16,10] 
IA_accept_party[2,4] <- state_stats[16,11] 
IA_accept_party[3,4] <- state_stats[16,12] 
IA_accept_party[4,4] <- state_stats[16,13] 
IA_accept_party[5,4] <- state_stats[16,5] 

IA_accept_party$Frequency <- 100*IA_accept_party$Count/IA_accept_party[5,2]
IA_accept_party$Rate <- 100*IA_accept_party$Count/IA_accept_party$Count2

colnames(IA_accept_party) <- c("Party", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")

IA_stats_requested <- IA_stats %>%
  select(County, Reg.Voters, Mail.Req.Tot, Pct.Request)

IA_stats_accepted <- IA_stats %>%
  select(County, Mail.Req.Tot, Mail.Accept.Tot, Pct.Accept)
```
## {.tabset}
Last Report: `r state_stats[16,9]`  
Source:  `r state_stats[16,2]`

### Returned and Accepted Mail Ballots

#### **Mail Ballots Returned and Accepted by Party Registration**

Iowa voters have cast **`r format(as.numeric(state_stats[16,6]), big.mark =",")`** mail ballots.

``` {r echo = FALSE}
kable(IA_accept_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

``` {r echo = FALSE}

IA_2020g_map_data <- IA_stats

IA_2020g_map_data$fips <- as.character(IA_2020g_map_data$fips)

IA_2020g_map_data <- mutate(IA_2020g_map_data, percent = round(100*Pct.Accept, digits = 1))

mapfile <- download_map_data("countries/us/us-ia-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

IA_2020g_map_data <- left_join(IA_2020g_map_data, mapdata, by = "fips")

IA_2020g_map_data <- arrange(IA_2020g_map_data, row)

hcmap(map = "countries/us/us-ia-all", data = IA_2020g_map_data,
  value = "percent", name = "Percent Accepted", joinBy = "fips") %>%
  hc_title(text ="Mail Ballot Return Rates") %>%
  hc_subtitle(text = "County plots may not be shaded using the same scale")
```

``` {r echo = FALSE}
datatable(IA_stats_accepted, colnames = c("County", "Mail Ballots Requested", "Mail Ballots Accepted", "Percent Accepted"), rownames = F) %>%
  formatPercentage('Pct.Accept', 1) %>%
  formatRound(c('Mail.Req.Tot', 'Mail.Accept.Tot'), 0, mark = ",")
```

### Requested Mail Ballots

Iowa voters have requested **`r format(as.numeric(state_stats[16,5]), big.mark =",")`** mail ballots.

#### **Mail Ballots Requested by Party Registration**

``` {r echo = FALSE}
kable(IA_req_send_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

``` {r echo = FALSE}

IA_2020g_map_data <- IA_stats

IA_2020g_map_data$fips <- as.character(IA_2020g_map_data$fips)

IA_2020g_map_data <- mutate(IA_2020g_map_data, percent = round(100*Pct.Request, digits = 1))

mapfile <- download_map_data("countries/us/us-ia-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

IA_2020g_map_data <- left_join(IA_2020g_map_data, mapdata, by = "fips")

IA_2020g_map_data <- arrange(IA_2020g_map_data, row)

hcmap(map = "countries/us/us-ia-all", data = IA_2020g_map_data,
  value = "percent", name = "Percent Requested", joinBy = "fips") %>%
  hc_title(text ="Mail Ballot Request Rates") %>%
  hc_subtitle(text = "County plots may not be shaded using the same scale")

```

``` {r echo = FALSE}
datatable(IA_stats_requested, colnames = c("County", "Voter Registration", "Mail Ballots Requested", "Percent Requested"), rownames = F) %>%
  formatPercentage('Pct.Request', 1) %>%
  formatRound(c('Reg.Voters', 'Mail.Req.Tot'), 0, mark = ",")
```
