---
title: "Illinois Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(DT)
library(highcharter)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")
IL_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote_IL.csv")
IL_FIPS <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/IL_FIPS.csv", col_types = "cc")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

# Illinois

IL_req_send_tot <- data.frame(Total=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)

IL_req_send_tot[1,1] <- "TOTAL"
IL_req_send_tot[1,2] <- sum(state_stats[14,5]) 

IL_req_send_tot$Percent <- 100.0


IL_stats_requests <- select(IL_stats, County, Reg.Voters, By.Mail, Pct.Request)

IL_stats_returns <- IL_stats %>%
  select(County, By.Mail.Returned, By.Mail) %>%
  mutate(Pct.Return = By.Mail.Returned/By.Mail)

IL_stats_voted <- IL_stats %>%
  select(County, Early.Tot, Reg.Voters) %>%
  mutate(Pct.Voted = Early.Tot/Reg.Voters)
```
## {.tabset}
Last Report: `r state_stats[14,9]`  
Source:  `r state_stats[14,2]`

For the maps, data for the cities of Bloomington, Chicago, Danville, East St. Louis, Galesburg, and Rockford have been added to their respective counties. **I am investigating an issue about how these county data are created. State statistics are correct.**

### Total In-person Early and Returned Mail Ballots  
Illinois voters have cast **`r format(as.numeric(state_stats[14,8]), big.mark =",")`** early votes.

``` {r echo = FALSE}
IL_stats_county <- IL_stats

IL_City_Bloomington_data <- IL_stats_county %>%
  filter(County == "City of Bloomington")

IL_City_Chicago_data <- IL_stats_county %>%
  filter(County == "City of Chicago")

IL_City_Danville_data <- IL_stats_county %>%
  filter(County == "City of Danville")

IL_City_ESL_data <- IL_stats_county %>%
  filter(County == "City of East St Louis")

IL_City_Galesburg_data <- IL_stats_county %>%
  filter(County == "City of Galesburg")

IL_City_Rockford_data <- IL_stats_county %>%
  filter(County == "City of Rockford")

IL_stats_county <- IL_stats_county %>%
  filter(County != "City of Bloomington" & County != "City of Chicago" & County != "City of Danville" & County != "City of East St Louis" & County != "City of Galesburg" & County != "City of Rockford")


IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "McLean County" ~ Reg.Voters + IL_City_Bloomington_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "McLean County" ~ By.Mail + IL_City_Bloomington_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "McLean County" ~ By.Mail.Returned + IL_City_Bloomington_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "McLean County" ~ Early + IL_City_Bloomington_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "McLean County" ~ Grace + IL_City_Bloomington_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Cook County" ~ Reg.Voters + IL_City_Chicago_data$Reg.Voters,
                                 TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Cook County" ~ By.Mail + IL_City_Chicago_data$By.Mail,
                                TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Cook County" ~ By.Mail.Returned + IL_City_Chicago_data$By.Mail.Returned,
                                TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Cook County" ~ Early + IL_City_Chicago_data$Early,
                                TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Cook County" ~ Grace + IL_City_Chicago_data$Grace,
                           TRUE ~ Grace))
  
IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Vermilion County" ~ Reg.Voters + IL_City_Danville_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Vermilion County" ~ By.Mail + IL_City_Danville_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Vermilion County" ~ By.Mail.Returned + IL_City_Danville_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Vermilion County" ~ Early + IL_City_Danville_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Vermilion County" ~ Grace + IL_City_Danville_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "St. Clair County" ~ Reg.Voters + IL_City_ESL_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "St. Clair County" ~ By.Mail + IL_City_ESL_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "St. Clair County" ~ By.Mail.Returned + IL_City_ESL_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "St. Clair County" ~ Early + IL_City_ESL_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "St. Clair County" ~ Grace + IL_City_ESL_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Knox County" ~ Reg.Voters + IL_City_Galesburg_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Knox County" ~ By.Mail + IL_City_Galesburg_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Knox County" ~ By.Mail.Returned + IL_City_Galesburg_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Knox County" ~ Early + IL_City_Galesburg_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Knox County" ~ Grace + IL_City_Galesburg_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Winnebago County" ~ Reg.Voters + IL_City_Rockford_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Winnebago County" ~ By.Mail + IL_City_Rockford_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Winnebago County" ~ By.Mail.Returned + IL_City_Rockford_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Winnebago County" ~ Early + IL_City_Rockford_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Winnebago County" ~ Grace + IL_City_Rockford_data$Grace,
                           TRUE ~ Grace))

IL_map_data <- left_join(IL_stats_county, IL_FIPS, by = "County")

IL_map_data <- mutate(IL_map_data, percent = round(100*(Early.Tot/Reg.Voters), digits = 1))

mapfile <- download_map_data("countries/us/us-il-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

IL_map_data <- left_join(IL_map_data, mapdata, by = "fips")

IL_map_data <- arrange(IL_map_data, row)

hcmap(map = "countries/us/us-il-all", data = IL_map_data,
      value = "percent", name = "Percent Voted", joinby = c("fips", "FIPS")) 
```

``` {r echo = FALSE}
datatable(IL_stats_voted, colnames = c("County", "Total Early Vote", "Voter Registration", "Percent Voted"), rownames = F) %>%
  formatPercentage('Pct.Voted', 1) %>%
  formatRound(c('Early.Tot','Reg.Voters'), 0, mark = ",")
```

### In-person Early Votes
Illinois voters have cast **`r format(as.numeric(state_stats[14,7]), big.mark =",")`** in-person early votes.

### Returned Mail Ballots

Illinois voters have cast **`r format(as.numeric(state_stats[14,6]), big.mark =",")`** mail ballots.

``` {r echo = FALSE}
IL_stats_county <- IL_stats

IL_City_Bloomington_data <- IL_stats_county %>%
  filter(County == "City of Bloomington")

IL_City_Chicago_data <- IL_stats_county %>%
  filter(County == "City of Chicago")

IL_City_Danville_data <- IL_stats_county %>%
  filter(County == "City of Danville")

IL_City_ESL_data <- IL_stats_county %>%
  filter(County == "City of East St Louis")

IL_City_Galesburg_data <- IL_stats_county %>%
  filter(County == "City of Galesburg")

IL_City_Rockford_data <- IL_stats_county %>%
  filter(County == "City of Rockford")

IL_stats_county <- IL_stats_county %>%
  filter(County != "City of Bloomington" & County != "City of Chicago" & County != "City of Danville" & County != "City of East St Louis" & County != "City of Galesburg" & County != "City of Rockford")


IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "McLean County" ~ Reg.Voters + IL_City_Bloomington_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "McLean County" ~ By.Mail + IL_City_Bloomington_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "McLean County" ~ By.Mail.Returned + IL_City_Bloomington_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "McLean County" ~ Early + IL_City_Bloomington_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "McLean County" ~ Grace + IL_City_Bloomington_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Cook County" ~ Reg.Voters + IL_City_Chicago_data$Reg.Voters,
                                 TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Cook County" ~ By.Mail + IL_City_Chicago_data$By.Mail,
                                TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Cook County" ~ By.Mail.Returned + IL_City_Chicago_data$By.Mail.Returned,
                                TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Cook County" ~ Early + IL_City_Chicago_data$Early,
                                TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Cook County" ~ Grace + IL_City_Chicago_data$Grace,
                           TRUE ~ Grace))
  
IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Vermilion County" ~ Reg.Voters + IL_City_Danville_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Vermilion County" ~ By.Mail + IL_City_Danville_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Vermilion County" ~ By.Mail.Returned + IL_City_Danville_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Vermilion County" ~ Early + IL_City_Danville_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Vermilion County" ~ Grace + IL_City_Danville_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "St. Clair County" ~ Reg.Voters + IL_City_ESL_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "St. Clair County" ~ By.Mail + IL_City_ESL_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "St. Clair County" ~ By.Mail.Returned + IL_City_ESL_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "St. Clair County" ~ Early + IL_City_ESL_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "St. Clair County" ~ Grace + IL_City_ESL_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Knox County" ~ Reg.Voters + IL_City_Galesburg_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Knox County" ~ By.Mail + IL_City_Galesburg_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Knox County" ~ By.Mail.Returned + IL_City_Galesburg_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Knox County" ~ Early + IL_City_Galesburg_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Knox County" ~ Grace + IL_City_Galesburg_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Winnebago County" ~ Reg.Voters + IL_City_Rockford_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Winnebago County" ~ By.Mail + IL_City_Rockford_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Winnebago County" ~ By.Mail.Returned + IL_City_Rockford_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Winnebago County" ~ Early + IL_City_Rockford_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Winnebago County" ~ Grace + IL_City_Rockford_data$Grace,
                           TRUE ~ Grace))

IL_map_data <- left_join(IL_stats_county, IL_FIPS, by = "County")

IL_map_data <- mutate(IL_map_data, percent = round(100*(By.Mail.Returned/By.Mail), digits = 1))

mapfile <- download_map_data("countries/us/us-il-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

IL_map_data <- left_join(IL_map_data, mapdata, by = "fips")

IL_map_data <- arrange(IL_map_data, row)

hcmap(map = "countries/us/us-il-all", data = IL_map_data,
      value = "percent", name = "Percent Returned", joinby = c("fips", "FIPS")) 
```

``` {r echo = FALSE}
datatable(IL_stats_returns, colnames = c("County", "Mail Ballots Returned", "Mail Ballot Requests", "Percent Returned"), rownames = F) %>%
  formatPercentage('Pct.Return', 1) %>%
  formatRound(c('By.Mail.Returned','By.Mail'), 0, mark = ",")
```

### Mail Ballot Requests

Illinois voters have requested **`r format(as.numeric(state_stats[14,5]), big.mark =",")`** mail ballots.

``` {r echo = FALSE}
IL_stats_county <- IL_stats

IL_City_Bloomington_data <- IL_stats_county %>%
  filter(County == "City of Bloomington")

IL_City_Chicago_data <- IL_stats_county %>%
  filter(County == "City of Chicago")

IL_City_Danville_data <- IL_stats_county %>%
  filter(County == "City of Danville")

IL_City_ESL_data <- IL_stats_county %>%
  filter(County == "City of East St Louis")

IL_City_Galesburg_data <- IL_stats_county %>%
  filter(County == "City of Galesburg")

IL_City_Rockford_data <- IL_stats_county %>%
  filter(County == "City of Rockford")

IL_stats_county <- IL_stats_county %>%
  filter(County != "City of Bloomington" & County != "City of Chicago" & County != "City of Danville" & County != "City of East St Louis" & County != "City of Galesburg" & County != "City of Rockford")


IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "McLean County" ~ Reg.Voters + IL_City_Bloomington_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "McLean County" ~ By.Mail + IL_City_Bloomington_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "McLean County" ~ By.Mail.Returned + IL_City_Bloomington_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "McLean County" ~ Early + IL_City_Bloomington_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "McLean County" ~ Grace + IL_City_Bloomington_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Cook County" ~ Reg.Voters + IL_City_Chicago_data$Reg.Voters,
                                 TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Cook County" ~ By.Mail + IL_City_Chicago_data$By.Mail,
                                TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Cook County" ~ By.Mail.Returned + IL_City_Chicago_data$By.Mail.Returned,
                                TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Cook County" ~ Early + IL_City_Chicago_data$Early,
                                TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Cook County" ~ Grace + IL_City_Chicago_data$Grace,
                           TRUE ~ Grace))
  
IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Vermilion County" ~ Reg.Voters + IL_City_Danville_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Vermilion County" ~ By.Mail + IL_City_Danville_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Vermilion County" ~ By.Mail.Returned + IL_City_Danville_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Vermilion County" ~ Early + IL_City_Danville_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Vermilion County" ~ Grace + IL_City_Danville_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "St. Clair County" ~ Reg.Voters + IL_City_ESL_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "St. Clair County" ~ By.Mail + IL_City_ESL_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "St. Clair County" ~ By.Mail.Returned + IL_City_ESL_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "St. Clair County" ~ Early + IL_City_ESL_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "St. Clair County" ~ Grace + IL_City_ESL_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Knox County" ~ Reg.Voters + IL_City_Galesburg_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Knox County" ~ By.Mail + IL_City_Galesburg_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Knox County" ~ By.Mail.Returned + IL_City_Galesburg_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Knox County" ~ Early + IL_City_Galesburg_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Knox County" ~ Grace + IL_City_Galesburg_data$Grace,
                           TRUE ~ Grace))

IL_stats_county <- IL_stats_county %>%
  mutate(Reg.Voters = case_when(County == "Winnebago County" ~ Reg.Voters + IL_City_Rockford_data$Reg.Voters,
                                TRUE ~ Reg.Voters)) %>%
  mutate(By.Mail = case_when(County == "Winnebago County" ~ By.Mail + IL_City_Rockford_data$By.Mail,
                             TRUE ~ By.Mail)) %>%
  mutate(By.Mail.Returned = case_when(County == "Winnebago County" ~ By.Mail.Returned + IL_City_Rockford_data$By.Mail.Returned,
                                      TRUE ~ By.Mail.Returned)) %>%
  mutate(Early = case_when(County == "Winnebago County" ~ Early + IL_City_Rockford_data$Early,
                           TRUE ~ Early)) %>%
  mutate(Grace = case_when(County == "Winnebago County" ~ Grace + IL_City_Rockford_data$Grace,
                           TRUE ~ Grace))

IL_map_data <- left_join(IL_stats_county, IL_FIPS, by = "County")

IL_map_data <- mutate(IL_map_data, percent = round(100*(By.Mail/Reg.Voters), digits = 1))

mapfile <- download_map_data("countries/us/us-il-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

IL_map_data <- left_join(IL_map_data, mapdata, by = "fips")

IL_map_data <- arrange(IL_map_data, row)

hcmap(map = "countries/us/us-il-all", data = IL_map_data,
      value = "percent", name = "Percent Requested", joinby = c("fips", "FIPS")) 
```

``` {r echo = FALSE}
datatable(IL_stats_requests, colnames = c("County", "Registered Voters", "Mail Ballot Requests", "Percent Requested"), rownames = F) %>%
  formatPercentage('Pct.Request', 1) %>%
  formatRound(c('Reg.Voters','By.Mail'), 0, mark = ",")
```
