---
title: "Maryland Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(DT)
library(highcharter)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")
MD_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote_MD.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"


party_shell_returned <- data.frame(Party=character(),
                                  Count=integer(),
                                  Frequency=double(),
                                  Count2=integer(),
                                  Rate=integer(),
                                  stringsAsFactors=FALSE)
party_shell_returned[1,1] <- "Democrats"
party_shell_returned[2,1] <- "Republicans"
party_shell_returned[3,1] <- "Minor"
party_shell_returned[4,1] <- "No Party Affiliation"
party_shell_returned[5,1] <- "TOTAL"


race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

# Maryland

MD_req_send_party <- party_shell

MD_req_send_party[1,2] <- state_stats[21,10] 
MD_req_send_party[2,2] <- state_stats[21,11] 
MD_req_send_party[3,2] <- state_stats[21,12] 
MD_req_send_party[4,2] <- state_stats[21,13] 
MD_req_send_party[5,2] <- state_stats[21,5] 

MD_req_send_party$Percent <- 100*MD_req_send_party$Count/MD_req_send_party[5,2]

MD_accept_party <- party_shell_returned

MD_accept_party[1,2] <- state_stats[21,29] 
MD_accept_party[2,2] <- state_stats[21,30] 
MD_accept_party[3,2] <- state_stats[21,31] 
MD_accept_party[4,2] <- state_stats[21,32] 
MD_accept_party[5,2] <- state_stats[21,6] 

MD_accept_party[1,4] <- state_stats[21,10] 
MD_accept_party[2,4] <- state_stats[21,11] 
MD_accept_party[3,4] <- state_stats[21,12] 
MD_accept_party[4,4] <- state_stats[21,13] 
MD_accept_party[5,4] <- state_stats[21,5] 

MD_accept_party$Frequency <- 100*MD_accept_party$Count/MD_accept_party[5,2]
MD_accept_party$Rate <- 100*MD_accept_party$Count/MD_accept_party$Count2

colnames(MD_accept_party) <- c("Party", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")

MD_inperson_party <- party_shell

MD_inperson_party[1,2] <- state_stats[21,77] 
MD_inperson_party[2,2] <- state_stats[21,78] 
MD_inperson_party[3,2] <- state_stats[21,79] 
MD_inperson_party[4,2] <- state_stats[21,80] 
MD_inperson_party[5,2] <- state_stats[21,7] 

MD_inperson_party$Percent <- 100*MD_inperson_party$Count/MD_inperson_party[5,2]

MD_voted_party <- party_shell

MD_voted_party[1,2] <- state_stats[21,77] + state_stats[21,29]
MD_voted_party[2,2] <- state_stats[21,78] + state_stats[21,30] 
MD_voted_party[3,2] <- state_stats[21,79] + state_stats[21,31] 
MD_voted_party[4,2] <- state_stats[21,80] + state_stats[21,32] 
MD_voted_party[5,2] <- state_stats[21,7] + state_stats[21,6]

MD_voted_party$Percent <- 100*MD_voted_party$Count/MD_voted_party[5,2]


MD_stats <- mutate(MD_stats, Pct.Request = Mail.Req.Tot/Reg.Voters)
MD_stats <- mutate(MD_stats, Pct.Accept = Mail.Rtn.Tot/Mail.Req.Tot)

MD_stats_requests <- select(MD_stats, County, Reg.Voters, Mail.Req.Tot, Pct.Request)
MD_stats_returns <- select(MD_stats, County, Mail.Req.Tot, Mail.Rtn.Tot, Pct.Accept)
```
## {.tabset}
Last Report: `r state_stats[21,9]`  
Source:  `r state_stats[21,2]`

Maryland reports mail ballots for all minor party and no-party affiliation voters in a single "Other category." I report these voters in the no-party affiliation category.

Maryland provides party registration statistics for mail ballots sent to voters, not requested by voters.

### Total Voted
Total Voted: **`r format(as.numeric(state_stats[21,8]), big.mark =",")`**

#### **Total Voted by Party Registration**
``` {r echo = FALSE}
kable(MD_voted_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

### In-Person Votes
In-Person Votes: **`r format(as.numeric(state_stats[21,7]), big.mark =",")`**

#### **In-Person Votes by Party Registration**
``` {r echo = FALSE}
kable(MD_inperson_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```


### Mail Ballots Returned

Mail Ballots Returned: **`r format(as.numeric(state_stats[21,8]), big.mark =",")`**

#### **Mail Ballots Returned and Accepted by Party Registration**

``` {r echo = FALSE}
kable(MD_accept_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

``` {r echo = FALSE}
MD_map_data <- MD_stats

MD_map_data <- mutate(MD_map_data, percent = round(100*(Mail.Rtn.Tot/Mail.Req.Tot), digits = 1))

# MD_map_data <- rename(MD_map_data, fips = FIPS_CODE)
MD_map_data <- mutate(MD_map_data, fips = as.character(fips))

mapfile <- download_map_data("countries/us/us-md-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

MD_map_data <- left_join(MD_map_data, mapdata, by = "fips")

MD_map_data <- arrange(MD_map_data, row)

hcmap(map = "countries/us/us-md-all", data = MD_map_data,
      value = "percent", name = "Percent Accepted", joinBy = "fips") %>%
  hc_title(text ="Mail Ballot Return Rates") %>%
  hc_subtitle(text = "County plots may not be shaded using the same scale")
```


``` {r echo = FALSE}
datatable(MD_stats_returns, colnames = c("County", "Mail Ballots Sent", "Mail Ballots Returned", "Percent Returned"), rownames = F) %>%
  formatPercentage('Pct.Accept', 1) %>%
  formatRound(c('Mail.Req.Tot', 'Mail.Rtn.Tot'), 0, mark = ",")
```

### Sent Mail Ballots

Ballots Sent: **`r format(as.numeric(state_stats[21,5]), big.mark =",")`**

#### **Mail Ballots Sent by Party Registration**

``` {r echo = FALSE}
kable(MD_req_send_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

``` {r echo = FALSE}
MD_map_data <-MD_stats

MD_map_data <- mutate(MD_map_data, percent = round(100*(Mail.Req.Tot/Reg.Voters), digits = 1))

# MD_map_data <- rename(MD_map_data, fips = FIPS_CODE)
MD_map_data <- mutate(MD_map_data, fips = as.character(fips))

mapfile <- download_map_data("countries/us/us-md-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

MD_map_data <- left_join(MD_map_data, mapdata, by = "fips")

MD_map_data <- arrange(MD_map_data, row)

hcmap(map = "countries/us/us-md-all", data = MD_map_data,
      value = "percent", name = "Percent Requested", joinBy = "fips") %>%
  hc_title(text ="Mail Ballot Sent Rates") %>%
  hc_subtitle(text = "County plots may not be shaded using the same scale")
```



``` {r echo = FALSE}
datatable(MD_stats_requests, colnames = c("County", "Registered Voters", "Mail Ballots Sent", "Percent Requested"), rownames = F) %>%
  formatPercentage('Pct.Request', 1) %>%
  formatRound(c('Reg.Voters','Mail.Req.Tot'), 0, mark = ",")
```

