---
title: "Michigan Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(DT)
library(highcharter)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")
MI_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote_MI.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

# Michigan

MI_req_send_tot <- data.frame(Total=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)

MI_req_send_tot[1,1] <- "TOTAL"
MI_req_send_tot[1,2] <- sum(state_stats[23,5]) 

MI_req_send_tot$Percent <- 100.0

MI_stats_requested <- MI_stats %>%
  select(CountyName, Reg.Voters, Mail.Req.Tot, Pct.Request)

MI_stats_accepted <- MI_stats %>%
  select(CountyName, Mail.Req.Tot, Mail.Accept.Tot, Pct.Accept)

```
## {.tabset}
Last Report: `r state_stats[23,9]`  
Source:  `r state_stats[23,2]`

These Michigan data are compiled from a file of absentee ballot requests purchased from the Michigan Secretary of State's office.

### Returned and Accepted Mail Ballots

Michigan voters have cast **`r format(as.numeric(state_stats[23,6]), big.mark =",")`** mail ballots.

``` {r echo = FALSE}

MI_2020g_map_data <- MI_stats

MI_2020g_map_data$fips <- as.character(MI_2020g_map_data$FIPS_NUM)

MI_2020g_map_data <- mutate(MI_2020g_map_data, percent = round(100*Pct.Accept, digits = 1))

mapfile <- download_map_data("countries/us/us-mi-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

MI_2020g_map_data <- left_join(MI_2020g_map_data, mapdata, by = "fips")

MI_2020g_map_data <- arrange(MI_2020g_map_data, row)

hcmap(map = "countries/us/us-mi-all", data = MI_2020g_map_data,
  value = "percent", name = "Percent Accepted", joinBy = "fips") 
```

``` {r echo = FALSE}
datatable(MI_stats_accepted, colnames = c("County", "Mail Ballots Requested", "Mail Ballots Accepted", "Percent Accepted"), rownames = F) %>%
  formatPercentage('Pct.Accept', 1) %>%
  formatRound(c('Mail.Req.Tot', 'Mail.Accept.Tot'), 0, mark = ",")
```

### Requested Mail Ballots

Michigan voters have requested **`r format(as.numeric(state_stats[23,5]), big.mark =",")`** mail ballots.

``` {r echo = FALSE}

MI_2020g_map_data <- MI_stats

MI_2020g_map_data$fips <- as.character(MI_2020g_map_data$FIPS_NUM)

MI_2020g_map_data <- mutate(MI_2020g_map_data, percent = round(100*Pct.Request, digits = 1))

mapfile <- download_map_data("countries/us/us-mi-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

MI_2020g_map_data <- left_join(MI_2020g_map_data, mapdata, by = "fips")

MI_2020g_map_data <- arrange(MI_2020g_map_data, row)

hcmap(map = "countries/us/us-mi-all", data = MI_2020g_map_data,
  value = "percent", name = "Percent Requested", joinBy = "fips") 
```

``` {r echo = FALSE}
datatable(MI_stats_requested, colnames = c("County", "Voter Registration", "Mail Ballots Requested", "Percent Requested"), rownames = F) %>%
  formatPercentage('Pct.Request', 1) %>%
  formatRound(c('Reg.Voters', 'Mail.Req.Tot'), 0, mark = ",")
```
