---
title: "Minnesota Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(DT)
library(highcharter)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")
MN_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote_MN.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

# Minnesota

MN_stats_returns <- select(MN_stats, County, Mail.Accept.Tot, Mail.Req.Tot, Pct.Accept)
MN_stats_requests <- select(MN_stats, County, Mail.Req.Tot, Reg.Voters, Pct.Request)
```
## {.tabset}

Last Report: `r state_stats[24,9]`  
Source:  `r state_stats[24,2]`  

### Total Early Vote  

Total Early Vote: **`r format(as.numeric(state_stats[24,8]), big.mark =",")`**  

Minnesota does not distinguish between mail and in-person ballots on their state reports. The statistics reported here thus combine all in-person early and mail ballot votes. 

``` {r echo = FALSE}
MN_map_data <- MN_stats

MN_map_data <- mutate(MN_map_data, percent = round(100*(Mail.Accept.Tot/Mail.Req.Tot), digits = 1))

MN_map_data <- mutate(MN_map_data, fips = as.character(fips))

mapfile <- download_map_data("countries/us/us-mn-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

MN_map_data <- left_join(MN_map_data, mapdata, by = "fips")

MN_map_data <- arrange(MN_map_data, row)

hcmap(map = "countries/us/us-mn-all", data = MN_map_data,
      value = "percent", name = "Percent Voted", joinBy = "fips") %>%
  hc_title(text ="Percent Accepted") %>%
  hc_subtitle(text = "County plots may not be shaded using the same scale")
```

``` {r echo = FALSE}
datatable(MN_stats_returns, colnames = c("County", "Total Early Vote", "Mail Ballots Requested", "Percent Voted"), rownames = F) %>%
  formatPercentage('Pct.Accept', 1) %>%
  formatRound(c('Mail.Accept.Tot','Mail.Req.Tot'), 0, mark = ",")
```

### Mail Ballots Requested  

Mail Ballots Requested: **`r format(as.numeric(state_stats[24,5]), big.mark =",")`**  

``` {r echo = FALSE}
MN_map_data <- MN_stats

MN_map_data <- mutate(MN_map_data, percent = round(100*(Mail.Req.Tot/Reg.Voters), digits = 1))

MN_map_data <- mutate(MN_map_data, fips = as.character(fips))

mapfile <- download_map_data("countries/us/us-mn-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

MN_map_data <- left_join(MN_map_data, mapdata, by = "fips")

MN_map_data <- arrange(MN_map_data, row)

hcmap(map = "countries/us/us-mn-all", data = MN_map_data,
      value = "percent", name = "Request Rate", joinBy = "fips") %>%
  hc_title(text ="Request Rate") %>%
  hc_subtitle(text = "County plots may not be shaded using the same scale")
```

``` {r echo = FALSE}
datatable(MN_stats_requests, colnames = c("County", "Mail Ballots Requested", "Registered Voters", "Percent Requested"), rownames = F) %>%
  formatPercentage('Pct.Request', 1) %>%
  formatRound(c('Mail.Req.Tot', 'Reg.Voters'), 0, mark = ",")
```
