---
title: "North Carolina Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(DT)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")

NC_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote_NC.csv")

# NC_stats <- NC_stats %>%
#  rename("Registered Voters" = Reg.Voters) %>%
#  rename("Mail Ballot Requests" = Mail.Ballot.Requests) %>%
#  rename("Percent Requested" = Pct.Requests)

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

# North Carolina

NC_req_send_party <- party_shell

NC_req_send_party[1,2] <- state_stats[34,10] 
NC_req_send_party[2,2] <- state_stats[34,11] 
NC_req_send_party[3,2] <- state_stats[34,12] 
NC_req_send_party[4,2] <- state_stats[34,13] 
NC_req_send_party[5,2] <- state_stats[34,5] 

NC_req_send_party$Percent <- 100*NC_req_send_party$Count/NC_req_send_party[5,2]

NC_req_send_race <- race_shell

NC_req_send_race[1,2] <- sum(state_stats[34,14])
NC_req_send_race[2,2] <- sum(state_stats[34,15]) 
NC_req_send_race[3,2] <- sum(state_stats[34,16]) 
NC_req_send_race[4,2] <- sum(state_stats[34,17]) 
NC_req_send_race[5,2] <- sum(state_stats[34,18]) 
NC_req_send_race[6,2] <- sum(state_stats[34,19]) 
NC_req_send_race[7,2] <- sum(state_stats[34,5]) 

NC_req_send_race$Percent <- 100 * NC_req_send_race$Count/NC_req_send_race[7,2]

NC_req_send_gender <- gender_shell

NC_req_send_gender[1,2] <- sum(state_stats[34,21])
NC_req_send_gender[2,2] <- sum(state_stats[34,20]) 
NC_req_send_gender[3,2] <- sum(state_stats[34,22]) 
NC_req_send_gender[4,2] <- sum(state_stats[34,5]) 

NC_req_send_gender$Percent <- 100 * NC_req_send_gender$Count/NC_req_send_gender[4,2]
```

Last Report: `r state_stats[34,9]`  
Source:  `r state_stats[34,2]`

### Discussion

North Carolina is the first state to begin general election voting. Election officials start sending ballots to voters on September 4.

### Mail Ballot Requests by Party Registration

``` {r echo = FALSE}
kable(NC_req_send_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

North Carolina registered Democrats have a **`r format(as.numeric(NC_req_send_party[1,2]-NC_req_send_party[2,2]), big.mark =",")`** ballot request lead over registered Republicans.

### Mail Ballot Requests by Race and Ethnicity

``` {r echo = FALSE}
kable(NC_req_send_race, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

### Mail Ballot Requests by Gender

``` {r echo = FALSE}
kable(NC_req_send_gender, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

### Mail Ballot Requests by County
North Carolina releases an update to the voter registration file each Saturday. The voter registration numbers are from the most recent release and may be slightly dated.

``` {r echo = FALSE}
#kable(NC_stats, format.args = list(big.mark = ",", 
#  scientific = FALSE), digits = 1) %>%
#  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

datatable(NC_stats, colnames = c("County", "Registered Voters", "Mail Ballot Requests", "Percent Requested"), rownames = F) %>%
  formatPercentage('Pct.Request', 1) %>%
  formatRound(c('Reg.Voters','Mail.Ballot.Requests'), 0, mark = ",")
```
