---
title: "Nevada Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(DT)
library(highcharter)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "No Party Affiliation/Minr"
party_shell[4,1] <- "TOTAL"

race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


party_shell_returned <- data.frame(Party=character(),
                                  Count=integer(),
                                  Frequency=double(),
                                  Count2=integer(),
                                  Rate=integer(),
                                  stringsAsFactors=FALSE)
party_shell_returned[1,1] <- "Democrats"
party_shell_returned[2,1] <- "Republicans"
party_shell_returned[3,1] <- "No Party Affiliation/Minor"
party_shell_returned[4,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

# Nevada

NV_inperson_party <- party_shell

NV_inperson_party[1,2] <- state_stats[29,77] 
NV_inperson_party[2,2] <- state_stats[29,78] 
NV_inperson_party[3,2] <- state_stats[29,80] 
NV_inperson_party[4,2] <- state_stats[29,7] 

NV_inperson_party$Percent <- 100*NV_inperson_party$Count/NV_inperson_party[4,2]


NV_voted_party <- party_shell

NV_voted_party[1,2] <- state_stats[29,77] + state_stats[29,29] 
NV_voted_party[2,2] <- state_stats[29,78] + state_stats[29,30] 
NV_voted_party[3,2] <- state_stats[29,80] + state_stats[29,32]
NV_voted_party[4,2] <- state_stats[29,7] + state_stats[29,6]

NV_voted_party$Percent <- 100*NV_voted_party$Count/NV_voted_party[4,2]



NV_req_send_party <- party_shell

NV_req_send_party[1,2] <- state_stats[29,10] 
NV_req_send_party[2,2] <- state_stats[29,11] 
NV_req_send_party[3,2] <- state_stats[29,13] 
NV_req_send_party[4,2] <- state_stats[29,5] 

NV_req_send_party$Percent <- 100*NV_req_send_party$Count/NV_req_send_party[4,2]

NV_accept_party <- party_shell_returned

NV_accept_party[1,2] <- state_stats[29,29] 
NV_accept_party[2,2] <- state_stats[29,30] 
NV_accept_party[3,2] <- state_stats[29,32] 
NV_accept_party[4,2] <- state_stats[29,6] 

NV_accept_party[1,4] <- state_stats[29,10] 
NV_accept_party[2,4] <- state_stats[29,11] 
NV_accept_party[3,4] <- state_stats[29,13] 
NV_accept_party[4,4] <- state_stats[29,5] 

NV_accept_party$Frequency <- 100*NV_accept_party$Count/NV_accept_party[4,2]
NV_accept_party$Rate <- 100*NV_accept_party$Count/NV_accept_party$Count2

colnames(NV_accept_party) <- c("Party", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")


```
## {.tabset}
Last Report: `r state_stats[29,9]`  
Source:  `r state_stats[29,2]`

**NOTE: The Nevada Secretary of State combines statistics for voters with no party affiliation and all minor parties. Nevada is including "curable" rejected mail ballots in its total mail ballot return statistics. I exclude these ballots.**  

### Total Voted

Total Voted: **`r format(as.numeric(state_stats[29,8]), big.mark =",")`**

#### **Total Voted by Party Registration**

``` {r echo = FALSE}
kable(NV_voted_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

### In-Person Votes

In-Person Votes: **`r format(as.numeric(state_stats[29,7]), big.mark =",")`**

#### **In-Person Votes by Party Registration**

``` {r echo = FALSE}
kable(NV_inperson_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```


### Returned Mail Ballots

Nevada voters have cast **`r format(as.numeric(state_stats[29,6]), big.mark =",")`** mail ballots.

#### **Mail Ballots Returned and Accepted by Party Registration**

``` {r echo = FALSE}
kable(NV_accept_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

### Mail Ballot Requests

Nevada voters have requested **`r format(as.numeric(state_stats[29,5]), big.mark =",")`** mail ballots.