---
title: "Ohio Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(DT)
library(highcharter)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")

OH_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote_OH.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

OH_req_send_tot <- data.frame(Total=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)

OH_req_send_tot[1,1] <- "TOTAL"
OH_req_send_tot[1,2] <- sum(state_stats[36,5]) 

OH_req_send_tot$Percent <- 100.0

OH_req_send_age <- age_shell

OH_req_send_age[1,2] <- state_stats[36,23] 
OH_req_send_age[2,2] <- state_stats[36,24] 
OH_req_send_age[3,2] <- state_stats[36,25] 
OH_req_send_age[4,2] <- state_stats[36,26] 
OH_req_send_age[5,2] <- state_stats[36,27] 
OH_req_send_age[6,2] <- state_stats[36,28] 
OH_req_send_age[7,2] <- state_stats[36,5] 

OH_req_send_age$Percent <- 100*OH_req_send_age$Count/OH_req_send_age[7,2]

OH_stats_requests <- select(OH_stats, COUNTY, Reg.Voters, Mail.Req.Tot, Pct.Request)
OH_stats_returns <- select(OH_stats, COUNTY, Mail.Rtn.Tot, Mail.Req.Tot, Pct.Return)
```
## {.tabset}
Last Report: `r state_stats[36,9]`  
Source:  `r state_stats[36,2]`

Data presented here are scraped from 80 of Ohio's 88 counties that post absentee ballot reports online. These 80 reporting counties include the largest counties of Cuyahoga, Franklin, and Hamilton.  

These data were last scraped on the Last Report date, but counties may have last posted their data on an earlier date.

You may see reports of Ohio mail ballot activity by "Party". I will not post party statistics because they are misleading. Ohio's party affiliation is unlike party registration in other states. In Ohio, "Party" is a record of the last party primary a voter participated in, in the current and past last two calendar years (see: [Ohio Code 3513.19(3)](https://codes.ohio.gov/orc/3513.19)). In Ohio's April 28, 2020 presidential primary, 124,628 more voters participated in the Democratic primary than the Republican primary, primarily because President Trump had only token opposition. As a consequence, **Ohio's "Party" is currently artificially tilted towards the Democrats.** Leveraging this quirk, both political parties and their allies have abused these party statistics over the years to make disingenuous claims about party enthusiasm in Ohio.

**I do not distinguish between mail and in-person votes in Ohio because some counties do not provide data by method of voting**  

### Mail Ballots Returned

Ballots Returned: **`r format(as.numeric(state_stats[36,6]), big.mark =",")`**

``` {r echo = FALSE}

OH_2020g_map_data <- OH_stats

OH_2020g_map_data$fips <- as.character(OH_2020g_map_data$fips)

OH_2020g_map_data <- mutate(OH_2020g_map_data, percent = round(100*Pct.Return, digits = 1))

mapfile <- download_map_data("countries/us/us-oh-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

OH_2020g_map_data <- left_join(OH_2020g_map_data, mapdata, by = "fips")

OH_2020g_map_data <- arrange(OH_2020g_map_data, row)

hcmap(map = "countries/us/us-oh-all", data = OH_2020g_map_data,
  value = "percent", name = "Percent Returned", joinBy = "fips") %>%
  hc_title(text ="Mail Ballot Return Rates") %>%
  hc_subtitle(text = "County plots may not be shaded using the same scale")
```

``` {r echo = FALSE}
datatable(OH_stats_returns, colnames = c("County", "Mail Ballots Returned", "Mail Ballot Requests", "Percent Returned"), rownames = F) %>%
  formatPercentage('Pct.Return', 1) %>%
  formatRound(c('Mail.Rtn.Tot','Mail.Req.Tot'), 0, mark = ",")
```

### Mail Ballots Requested

Ballots Requested: **`r format(as.numeric(state_stats[36,5]), big.mark =",")`**

#### Mail Ballot Requests by County
Ohio releases a weekly update to the voter registration file. The voter registration numbers are from the most recent release and may be slightly dated.

``` {r echo = FALSE}

OH_2020g_map_data <- OH_stats

OH_2020g_map_data$fips <- as.character(OH_2020g_map_data$fips)

OH_2020g_map_data <- mutate(OH_2020g_map_data, percent = round(100*Pct.Request, digits = 1))

mapfile <- download_map_data("countries/us/us-oh-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

OH_2020g_map_data <- left_join(OH_2020g_map_data, mapdata, by = "fips")

OH_2020g_map_data <- arrange(OH_2020g_map_data, row)

hcmap(map = "countries/us/us-oh-all", data = OH_2020g_map_data,
  value = "percent", name = "Percent Requested", joinBy = "fips") %>%
  hc_title(text ="Mail Ballot Return Rates") %>%
  hc_subtitle(text = "County plots may not be shaded using the same scale")
```

``` {r echo = FALSE}
datatable(OH_stats_requests, colnames = c("County", "Registered Voters", "Mail Ballot Requests", "Percent Requested"), rownames = F) %>%
  formatPercentage('Pct.Request', 1) %>%
  formatRound(c('Reg.Voters','Mail.Req.Tot'), 0, mark = ",")
```
