---
title: "Pennsylvania Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"

race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"

# Pennsylvania

PA_req_send_party <- party_shell

PA_req_send_party[1,2] <- state_stats[39,10] 
PA_req_send_party[2,2] <- state_stats[39,11] 
PA_req_send_party[3,2] <- state_stats[39,12] 
PA_req_send_party[4,2] <- state_stats[39,13] 
PA_req_send_party[5,2] <- state_stats[39,5] 

PA_req_send_party$Percent <- 100*PA_req_send_party$Count/PA_req_send_party[5,2]

PA_req_send_age <- age_shell

PA_req_send_age[1,2] <- state_stats[39,23] 
PA_req_send_age[2,2] <- state_stats[39,24] 
PA_req_send_age[3,2] <- state_stats[39,25] 
PA_req_send_age[4,2] <- state_stats[39,26] 
PA_req_send_age[5,2] <- state_stats[39,27] 
PA_req_send_age[6,2] <- state_stats[39,28] 
PA_req_send_age[7,2] <- state_stats[39,5] 

PA_req_send_age$Percent <- 100*PA_req_send_age$Count/PA_req_send_age[7,2]
```
Last Report: `r state_stats[39,9]`  
Source:  `r state_stats[39,2]`

### Discussion

Pennsylvania ballot requests are good through the end of a calendar year unless a voter otherwise cancels it. I analyze the ballot requests for the state's closed primary. The number of ballot requests from all voters, especially minor party and no party affiliated voters, is likely larger than reported here.

### Mail Ballot Requests by Party Registration

``` {r echo = FALSE}
kable(PA_req_send_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

Pennsylvania registered Democrats have a **`r format(as.numeric(PA_req_send_party[1,2]-PA_req_send_party[2,2]), big.mark =",")`** ballot request lead over registered Republicans.

### Mail Ballot Requests by Age

``` {r echo = FALSE}
kable(PA_req_send_age, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```


