---
title: "Pennsylvania Early Voting Statistics"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)

state_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote.csv")
PA_stats <- read_csv("D:/DropBox/Dropbox/Mail_Ballots_2020/markdown/2020G_Early_Vote_PA.csv")

# Setup

party_shell <- data.frame(Party=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
party_shell[1,1] <- "Democrats"
party_shell[2,1] <- "Republicans"
party_shell[3,1] <- "Minor"
party_shell[4,1] <- "No Party Affiliation"
party_shell[5,1] <- "TOTAL"


party_shell_returned <- data.frame(Party=character(),
                                  Count=integer(),
                                  Frequency=double(),
                                  Count2=integer(),
                                  Rate=integer(),
                                  stringsAsFactors=FALSE)
party_shell_returned[1,1] <- "Democrats"
party_shell_returned[2,1] <- "Republicans"
party_shell_returned[3,1] <- "Minor"
party_shell_returned[4,1] <- "No Party Affiliation"
party_shell_returned[5,1] <- "TOTAL"


race_shell <- data.frame(Race=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
race_shell[1,1] <- "Non-Hispanic White"
race_shell[2,1] <- "Non-Hispanic Black"
race_shell[3,1] <- "Hispanic"
race_shell[4,1] <- "Non-Hispanic Asian American"
race_shell[5,1] <- "Non-Hispanic Native American"
race_shell[6,1] <- "Other/Multiple/Unknown"
race_shell[7,1] <- "TOTAL"


gender_shell <- data.frame(Gender=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
gender_shell[1,1] <- "Female"
gender_shell[2,1] <- "Male"
gender_shell[3,1] <- "Unknown"
gender_shell[4,1] <- "TOTAL"

age_shell <- data.frame(Age=character(),
                                  Count=integer(),
                                  Percent=double(),
                                  stringsAsFactors=FALSE)
age_shell[1,1] <- "18 to 24"
age_shell[2,1] <- "25 to 34"
age_shell[3,1] <- "35 to 44"
age_shell[4,1] <- "45 to 54"
age_shell[5,1] <- "55 to 64"
age_shell[6,1] <- "65 and up"
age_shell[7,1] <- "TOTAL"


age_shell_returned <- data.frame(Race=character(),
                                  Count=integer(),
                                  Frequency=double(),
                                  Count2=integer(),
                                  Rate=integer(),
                                  stringsAsFactors=FALSE)
age_shell_returned[1,1] <- "18 to 24"
age_shell_returned[2,1] <- "25 to 34"
age_shell_returned[3,1] <- "35 to 44"
age_shell_returned[4,1] <- "45 and 55"
age_shell_returned[5,1] <- "56 and 65"
age_shell_returned[6,1] <- "66 and up"
age_shell_returned[7,1] <- "TOTAL"

# Pennsylvania

PA_req_send_party <- party_shell

PA_req_send_party[1,2] <- state_stats[39,10] 
PA_req_send_party[2,2] <- state_stats[39,11] 
PA_req_send_party[3,2] <- state_stats[39,12] 
PA_req_send_party[4,2] <- state_stats[39,13] 
PA_req_send_party[5,2] <- state_stats[39,5] 

PA_req_send_party$Percent <- 100*PA_req_send_party$Count/PA_req_send_party[5,2]


PA_accept_party <- party_shell_returned

PA_accept_party[1,2] <- state_stats[39,29] 
PA_accept_party[2,2] <- state_stats[39,30] 
PA_accept_party[3,2] <- state_stats[39,31] 
PA_accept_party[4,2] <- state_stats[39,32] 
PA_accept_party[5,2] <- state_stats[39,6] 

PA_accept_party[1,4] <- state_stats[39,10] 
PA_accept_party[2,4] <- state_stats[39,11] 
PA_accept_party[3,4] <- state_stats[39,12] 
PA_accept_party[4,4] <- state_stats[39,13] 
PA_accept_party[5,4] <- state_stats[39,5] 

PA_accept_party$Frequency <- 100*PA_accept_party$Count/PA_accept_party[5,2]
PA_accept_party$Rate <- 100*PA_accept_party$Count/PA_accept_party$Count2

colnames(PA_accept_party) <- c("Party", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")

PA_req_send_age <- age_shell

PA_req_send_age[1,2] <- state_stats[39,23] 
PA_req_send_age[2,2] <- state_stats[39,24] 
PA_req_send_age[3,2] <- state_stats[39,25] 
PA_req_send_age[4,2] <- state_stats[39,26] 
PA_req_send_age[5,2] <- state_stats[39,27] 
PA_req_send_age[6,2] <- state_stats[39,28] 
PA_req_send_age[7,2] <- state_stats[39,5] 

PA_req_send_age$Percent <- 100*PA_req_send_age$Count/PA_req_send_age[7,2]

PA_accept_age <- age_shell_returned

PA_accept_age[1,2] <- sum(state_stats[39,42])
PA_accept_age[2,2] <- sum(state_stats[39,43]) 
PA_accept_age[3,2] <- sum(state_stats[39,44]) 
PA_accept_age[4,2] <- sum(state_stats[39,45]) 
PA_accept_age[5,2] <- sum(state_stats[39,46])
PA_accept_age[6,2] <- sum(state_stats[39,47])
PA_accept_age[7,2] <- sum(state_stats[39,6])

PA_accept_age[1,4] <- sum(state_stats[39,23])
PA_accept_age[2,4] <- sum(state_stats[39,24]) 
PA_accept_age[3,4] <- sum(state_stats[39,25]) 
PA_accept_age[4,4] <- sum(state_stats[39,26]) 
PA_accept_age[5,4] <- sum(state_stats[39,27])
PA_accept_age[6,4] <- sum(state_stats[39,28])
PA_accept_age[7,4] <- sum(state_stats[39,5])

PA_accept_age$Frequency <- 100 * PA_accept_age$Count/PA_accept_age[7,2]
PA_accept_age$Rate <- 100*PA_accept_age$Count/PA_accept_age$Count2

colnames(PA_accept_age) <- c("Age", "Returned Ballots", "Freq. Distribution", "Requested Ballots", "Return Rate")

PA_stats_requests <- select(PA_stats, County, Reg.Voters, Mail.Req.Tot, Pct.Request)
PA_stats_returns <- select(PA_stats, County, Mail.Req.Tot, Mail.Rtn.Tot, Pct.Accept)
```
## {.tabset}  
Last Report: `r state_stats[39,9]`  
Source:  `r state_stats[39,2]`

### Mail Ballots Returned

#### **Mail Ballots Returned by Party Registration**

``` {r echo = FALSE}
kable(PA_accept_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

``` {r echo = FALSE}
PA_map_data <- PA_stats

PA_map_data <- mutate(PA_map_data, percent = round(100*(Mail.Rtn.Tot/Mail.Req.Tot), digits = 1))

PA_map_data <- mutate(PA_map_data, fips = as.character(fips))

mapfile <- download_map_data("countries/us/us-pa-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

PA_map_data <- left_join(PA_map_data, mapdata, by = "fips")

PA_map_data <- arrange(PA_map_data, row)

hcmap(map = "countries/us/us-pa-all", data = PA_map_data,
      value = "percent", name = "Percent Returned", joinBy = "fips") %>%
  hc_title(text ="Mail Ballot Return Rates") %>%
  hc_subtitle(text = "County plots may not be shaded using the same scale")
```

``` {r echo = FALSE}
datatable(PA_stats_returns, colnames = c("County", "Mail Ballots Requested", "Mail Ballots Returned", "Percent Returned"), rownames = F) %>%
  formatPercentage('Pct.Accept', 1) %>%
  formatRound(c('Mail.Req.Tot', 'Mail.Rtn.Tot'), 0, mark = ",")
```


#### **Mail Ballots Returned by Age**

``` {r echo = FALSE}
kable(PA_accept_age, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

### Mail Ballot Requests

#### Mail Ballot Requests by Party Registration

``` {r echo = FALSE}
kable(PA_req_send_party, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

Pennsylvania registered Democrats have a **`r format(as.numeric(PA_req_send_party[1,2]-PA_req_send_party[2,2]), big.mark =",")`** ballot request lead over registered Republicans.

``` {r echo = FALSE}
PA_map_data <- PA_stats

PA_map_data <- mutate(PA_map_data, percent = round(100*(Mail.Req.Tot/Reg.Voters), digits = 1))

PA_map_data <- mutate(PA_map_data, fips = as.character(fips))

mapfile <- download_map_data("countries/us/us-pa-all.js")
mapdata <- get_data_from_map(mapfile)

mapdata$row <- as.integer(rownames(mapdata))

PA_map_data <- left_join(PA_map_data, mapdata, by = "fips")

PA_map_data <- arrange(PA_map_data, row)

hcmap(map = "countries/us/us-pa-all", data = PA_map_data,
      value = "percent", name = "Percent Requested", joinBy = "fips") %>%
  hc_title(text ="Mail Ballot Request Rates") %>%
  hc_subtitle(text = "County plots may not be shaded using the same scale")
```

``` {r echo = FALSE}
datatable(PA_stats_requests, colnames = c("County", "Registered Voters", "Mail Ballots Requested", "Percent Requested"), rownames = F) %>%
  formatPercentage('Pct.Request', 1) %>%
  formatRound(c('Reg.Voters','Mail.Req.Tot'), 0, mark = ",")
```

#### Mail Ballot Requests by Age

``` {r echo = FALSE}
kable(PA_req_send_age, format.args = list(big.mark = ",", 
  scientific = FALSE), digits = 1) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

